Impaired central pattern generators due to abnormal EPHA4 signaling leads to idiopathic scoliosis

Version 1

Abstract

Abstract

Idiopathic scoliosis (IS) is the most common form of spinal deformity with unclear pathogenesis. In this study, we firstly reanalyzed the loci associated with IS, drawing upon previous studies. Subsequently, we mapped these loci to candidate genes using either location-based or function-based strategies. To further substantiate our findings, we verified the enrichment of variants within these candidate genes across several large IS cohorts encompassing Chinese, East Asian, and European populations. Consequently, we identified variants in the EPHA4 gene as compelling candidates for IS. To confirm their pathogenicity, we generated zebrafish mutants of epha4a. Remarkably, the zebrafish epha4a mutants exhibited pronounced scoliosis during later stages of development, effectively recapitulating the IS phenotype. We observed that the epha4a mutants displayed defects in left-right coordination during locomotion, which arose from disorganized neural activation in these mutants. Our subsequent experiments indicated that the disruption of the central pattern generator (CPG) network, characterized by abnormal axon guidance of spinal cord interneurons, contributed to the disorganization observed in the mutants. Moreover, when knocked down efnb3b, the ligand for Epha4a, we observed similar CPG defects and disrupted left-right locomotion. These findings strongly suggested that ephrin B3-Epha4 signaling is vital for the proper functioning of CPGs, and defects in this pathway could lead to scoliosis in zebrafish. Furthermore, we identified two cases of IS in NGEF, a downstream molecule in the EPHA4 pathway. Collectively, our data provide compelling evidence that neural patterning impairments and disruptions in CPGs may underlie the pathogenesis of IS.

Introduction

Idiopathic scoliosis (IS) is the most common form of spinal deformity, affecting about 2.5% of the global population (Hresko 2013, Luk et al. 2010). IS may have long-term physical and mental health consequences, such as cosmetic deformity, cardiopulmonary impairment, and even disability (Weinstein et al. 2003). All these consequences can severely reduce the quality of life. Early intervention with conservative treatments, such as braces, can control scoliosis progression and reduce the need for surgical intervention (Weinstein et al. 2013). However, IS often remains undiagnosed until malformation is evident, emphasizing the importance of risk-prediction measurements in medical management.

Genetic factors are thought to play a significant role in the development of IS, while only a few genes have been associated with the condition to date (Miller 2007, Cheng et al. 2015). Several single-nucleotide polymorphisms (SNPs) associated with susceptibility to IS have been identified through genome-wide association studies (GWASs), including SNPs linked to genes such as LBX1, GPR126, and BNC2. Notably, knockdown or overexpression of zebrafish homologs of these IS-associated genes have yielded body axis defects (Kou et al. 2019, Zhu et al. 2015, Guo et al. 2016, Kou et al. 2013, Ogura et al. 2015). In addition to common SNPs, rare variants with larger effect sizes or causing rare Mendelian disease traits also contribute to IS. Rare variants in FBN1, FBN2, and other extracellular matrix genes are associated with severe IS (Buchan et al. 2014, Haller et al. 2016). Linkage analysis of familial IS suggests that it follows autosomal dominant (AD), X-linked dominant, and multifactorial patterns of inheritance (Miller et al. 2005). Variants in CHD7 and AKAP2 are also implicated in the pathogenesis of Mendelian forms of IS (Gao et al. 2007, Li et al. 2016). Finally, centriolar protein POC5 and planar cell polarity protein Vang-like protein 1 (VANGL1) were also shown to be associated with IS. Notably, all these genes participate in a wide range of biological processes, the mechanisms underlying IS progression still require further investigation. Consequently, there is currently no consensus on the etiology of IS (Tang et al. 2021).

In this study, we developed a novel pipeline that combines SNP-to-gene mapping and rare variant association analysis. Using this approach, we analyzed a large Chinese population with IS and further validated our findings in East Asian and European populations. Through these analyses, we identified EPHA4 as a novel gene associated with IS. To confirm the pathogenicity of this gene, we used a zebrafish model and found that impaired EPHA4 pathway components and resulting defects in central pattern generators (CPGs) are essential factors in the pathogenesis of IS. Additionally, we searched for candidate genes related to EPHA4 signaling pathways and identified mutations in the NGEF among IS patients. Overall, our data suggest that the impairment of the EPHA4 pathway and central pattern generators plays a previously unknown role in the development and progression of IS.

Results

Enrichment analysis of rare variants in IS candidate genes

To identify candidate genes associated with IS, we conducted a comprehensive literature review of SNPs that are linked to IS from 14 published GWASs. This led us to obtain forty-one SNPs that showed genome-wide significance (P < 5×10−8) as detailed in Table S1. We then employed positional mapping, expression quantitative trait locus (eQTL) mapping, or chromatin interaction mapping to identify 156 candidate genes that could be potentially linked to these SNPs. Further analysis of rare variants for the 156 candidate genes using exome data from 411 Chinese probands and 3,800 unrelated Chinese controls identified EPHA4 as the only significant gene (P=0.045, OR=4.09) (Table S2). EPHA4 is a member of the Eph family of receptor tyrosine kinases which play a vital role in the development of the nervous system (Kania and Klein 2016). We identified three rare variants in EPHA4 from three IS patients, including one splicing-donor variant and two missense variants (Table S2). The flowchart of the entire pipeline is shown in Fig. S1.

Inheritance pattern and functional analyses of variants in EPHA4

We characterized the inheritance pattern of variants in EPHA4 using either trio sequencing data or Sanger sequencing for the parents. None of the parents being tested had scoliosis based on a clinical screening test. Of the three rare variants in EPHA4, two (NM_004438.3: c.1443+1G>C and c.2546G>A [p.Cys849Tyr]) arose de novo (Fig. 1A and Table 1). The heterozygous splicing variant (c.1443+1G>C) was identified in a 15-year-old female IS patient (SCO2003P0846) (Fig. 1B). Scoliosis in this patient was diagnosed at 13 years of age, with a main curve Cobb angle of 60°. The in vitro minigene assay showed that the c.1443+1G>C variant introduced a new splicing site, resulting in a 36-bp in-frame deletion in exon 6 (Fig. 1G, 1I, S2A). The EPHA4 heterozygous missense variant (c.2546G>A, p.Cys849Tyr), identified in a 13-year-old female IS patient (SCO2003P2146) (Fig. 1C), is located in the protein tyrosine kinase domain of EPHA4 protein (Fig. 1A). The onset of scoliosis in this patient was at 11 years of age with a 70° main curve Cobb angle. As EPHA4 plays a crucial role in the phosphorylation of CDK5, which subsequently activates the downstream pathways (Fu et al. 2007), we conducted western blotting to quantify the phosphorylation level of CDK5. Our results revealed that the missense variant resulted in a decreased phosphorylation level of CDK5 (p=0.015), suggesting a partial loss-of-function (LoF) of EPHA4 (Fig. 1F).

Clinical and genetic information on IS patients and functional effect of EPHA4 variants.

A: Locations of three EPHA4 single nucleotide variants relative to the protein domains (top panel) and exons 1-18 (bottom panel). B-E: Pedigrees and spinal radiographs of four probands with dominant gene variants. Sanger sequencing confirmed the variants. The arrows indicate the probands. The term +/+ denotes the wild-type, and cDNA change/+ denotes the heterozygous variant. F: Western blot analysis of EPHA4-c.2546G>A variant showing the protein expression levels of EPHA4 and CDK5 and the amount of phosphorylated CDK5 (pCDK5) expressed from the EPHA4-mutant plasmid. WT: wild type. G: Schematic representation of the effect of the EPHA4-c.1443+1G>C mutation on the splicing process. This variant induced a new splicing site (red box). The yellow box indicates the splicing acceptor. H: Schematic representation of the effect of the EPHA4-c.1318+10344A>G mutation on the splicing process. This variant induced a new splicing site (red box). The yellow box indicates the splicing acceptor. I: The minigene assay result showed that the c.1443+1G>C variant introduced a new splicing site, resulting in a 36-bp in-frame deletion in exon 6. J: The nested PCR showed that the c.1318+10344A>G variant induced exon 5 skipping, resulting in a 339-bp in-frame deletion. K: NCBI RefSeq genes included in 2q35-q36.2 from UCSC Genome Browser. EPHA4 is shown by the red oval.

We next investigated de novo non-coding variants of EPHA4 via whole genome sequencing (WGS) in 116 trio families with IS. A de novo heterozygous EPHA4 intronic variant (c.1318+10344A>G) was identified in a 16-year-old female patient (SCO2003P2080) (Fig. 1D and Table 1). She developed scoliosis at 11 years of age with a 70° main curve Cobb angle. This variant is predicted to affect the branch point of the fifth intron of EPHA4 (Fig. S2B). We performed nested PCR to show that this variant induced exon 5 skipping, resulting in a 339-bp in-frame deletion (Fig. 1H, 1J).

Next, we employed an in-house gene matching approach under the framework of DISCO study, which identified a 6-year-old patient (SCO2003P3202) who had been previously diagnosed with Waardenburg syndrome caused by a 4.46 Mb de novo deletion at 2q35-q36.2 (Li et al. 2015) (Table 1). Intriguingly, the patient also presented mild scoliosis (Fig. 1E). The deletion included the entire PAX3 gene, which was responsible for the Waardenburg syndrome phenotype, and 36 neighboring genes, including EPHA4 (Fig. 1K). As scoliosis is not typically associated with Waardenburg syndrome caused solely by PAX3 pathogenic variants (Tassabehji et al. 1993), we hypothesize that the deletion of EPHA4 may be responsible for the IS phenotype in this patient.

Notably, the GWAS signal which we mapped to EPHA4 (rs13398147) (Zhu et al. 2015) represents a significant eQTL in esophagus and colon tissues, with the T allele associated with decreased expression of EPHA4. In our East Asian GWAS cohort of 6,449 adolescent IS patients and 158,068 controls, we identified another two eQTLs in EPHA4 associated with decreased expression of EPHA4 in brain tissue (Table S3). In the same GWAS cohort, common SNPs in EPHA4, after aggregation, also showed significant enrichment (P=0.023) in IS patients versus controls. Taken together, the convergence between rare and common variants of EPHA4 that lead to LoF or hypomorphic effects highlights the pivotal role of EPHA4 in the pathogenesis of IS.

IS-like phenotypes in zebrafish epha4 mutants

To investigate the role of EPHA4 in scoliosis, we utilized zebrafish as a model system due to its versatile nature in modeling adolescent IS (Grimes et al. 2016, Bagnat and Gray 2020, Xie et al. 2022, Boswell and Ciruna 2017). Zebrafish have two homologs of EPHA4, epha4a and epha4b. Using CRISPR-Cas9, we established a stable epha4a zebrafish mutant line with a 63-bp deletion in exon 3, which introduced a stop codon and resulted in a truncated protein (Fig. 2A-B). The homozygous epha4a mutant larvae had no apparent defects in either notochord or body axis development (Fig. S3A). While in consistent with previous reports, the hindbrain rhombomeric boundaries were found to be defective in both the epha4a mutants and morphants (Fig. S3B) (Cayuso et al. 2019, Cooke, Kemp and Moens 2005). Interestingly, more than 75% adult mutants showed mild scoliosis (88 of 116), and some mutants exhibited severe scoliotic phenotype (4 of 116) (Fig. 2C-D, Movie S1&S2). Remarkably, some heterozygous adult mutants also developed mild scoliosis (14 of 95), whereas none of the wild-type fish showed any signs of scoliosis (0 of 76) (Fig. 2C-D). Similarly, we further generated the epha4b mutants with a 25-bp deletion in exon 3, which resulted in a frameshift mutation (Fig. S4A). The homozygous epha4b zebrafish mutants also developed mild scoliosis (28 of 43) (Fig. S4B-C). Intriguingly, both epha4a and epha4b mutants exhibited early onset scoliosis starting from around 20 days post fertilization (Fig. S4D), a stage similar to that of IS patients. Collectively, these data suggest that mutations in Epha4 proteins are linked to the scoliotic phenotype in zebrafish.

Scoliosis in zebrafish epha4a mutants.

A: Diagram of the protein domains, genomic structures, and sequences of wild-type and corresponding epha4a mutants. Red arrows indicate mutation sites. Blue boxes indicate open reading frames. Underlined sequences indicate the protospacer adjacent motif (PAM) region, and red fonts indicate Cas9 binding sites. LBD: ligand binding domain; FN3: fibronectin type 3 domain; TMD: transmembrane domain; PTKc: catalytic domain of the protein tyrosine kinases; SAM: sterile alpha motif. B: Sanger sequencing results confirmed the deletion of the target region in epha4a mutant transcripts. C: Bar graph showing the percentages of adult zebrafish with normal, mild, or severe body axis defects in wild-type (n=76), epha4a heterozygote (n=95), or epha4a mutants (n=116). D: Representative images of wild-type and epha4a mutants. Micro CT images are shown on the right. Lateral and dorsal views are shown. Scale bar: 1 cm.

Abnormal left-right swimming pattern in the absence of Epha4a

While the epha4a mutants seemed to be grossly normal during the larvae stages, they developed spinal curvature gradually during later development. We decided to investigate whether these mutants exhibited any abnormalities in their behavior at the larvae stages using EthoVision XT software. By monitoring the swimming behavior of 8 days post-fertilization (dpf) zebrafish larvae, we found the motion distance and swimming velocity were significantly decreased in epha4a mutants (Fig. 3A-B). In addition, we observed a remarkable difference in the relative turning angle and angular velocity between these two groups. The wild-type group changed their swimming direction randomly, showing a relative angle around 0 degrees (Fig. 3C-D). In contrast, the epha4a mutants favored turning to one side of their directions (Fig 3C-D). Moreover, the absolute turn angles and turning speed (angular velocity) were significantly higher in the mutants (Fig. 3E-F).

Abnormal left-right swimming pattern in the absence of Epha4a.

A: Dot plots showing the swimming distance of each 8 dpf larva at a duration of 4 min (N=54 for wild-type and N=60 for epha4a mutants). B: Dot plots showing the swimming velocity of wild-type and mutant larvae as indicated. C: Diagram showing the turning angle (θ) of the larvae during swimming. D: Scatter plot showing the relative turning angle of wild-type and mutant larvae. The relative angles were calculated by the sum of turning angles during fish swimming with left (positive) or right (negative) turns. The epha4a mutants favored turning to one side of their directions compared with those of wild-type larvae. E: Dot plots showing the average absolute turning angle of wild-type and mutant larvae as indicated. F: Dot plots showing the average absolute angular velocity of wild-type and mutant larvae as indicated. G: Bar graph showing the percentages of 5 dpf zebrafish larvae with different reactions after tactile stimulation. N=10 for each group; the numbers of tactile stimulations are indicated in each column. H: Representative images of the total body curvature measurements in zebrafish larvae, with values as the sum of α and β angles shown in the figure. I: Representative time-series images of 5 dpf wild-type and epha4a mutant zebrafish larvae after tactile stimulation to the head. Each panel represents the points of maximal body curvature for the C-bend and counterbend after the tactile startle response. J, K: A plot of body curvature angles as measured in panel (H) during swimming in response to tactile head stimulation in 5 dpf wild-type (N=5 larvae, n=10 stimuli) and epha4a mutants (N=5 larvae, n=7 stimuli). L: The maximum curvature angles during the first C-bend and counterbend after tactile stimulation in wild-type and epha4a mutant larvae.

Next, we compared the swimming behavior after startle response between wild-type and epha4a mutants. We used a needle to touch the head or the tail of 5 days post-fertilization (dpf) zebrafish larvae to stimulate swimming behavior. In sibling controls, the larvae responded to the tactile stimulation and swam away quickly (Movie S3). Conversely, the epha4a mutants failed to respond to the initial stimulation, and the swimming pattern was defective with an abnormal bending pattern (37 of 62 from tail-stimulation group and 34 of 77 from head-stimulation group) (Fig. 3G, Movie S4). We further performed more comprehensive analysis by high-speed video microscopy. After tactile stimulation, wild-type larvae displayed a high-speed C-bend turn followed by a weaker counterbend turn after several milliseconds (Fig. 3H-I). This rhythmic left/right swimming pattern ensures that the fish swim away from the frightening stimulus. The turning angles of the control larvae after stimulation changed with a sinusoidal wave pattern (Fig. 3J). In contrast, this pattern was dramatically different in epha4a mutants (Fig. 3I,K). Of note, although the C-bend turning angles were similar between epha4a mutants and control siblings, the turning angles of the counterbend decreased significantly (Fig. 3L), implying a left/right coordination defect. Taken together, both regular swimming and tactile stimulation analyses suggested that the left-right coordination swimming pattern is compromised in the absence of Epha4a.

Defects of left-right coordination due to abnormal CPG in the absence of EphA4

The coordinated left-right locomotion of zebrafish larvae relies on the synchronized contraction of muscle fibers, a process regulated by motor neurons situated on each side of the fish. To explore this intricate mechanism, we utilized a Tg(elavl3:GAL4; UAS: GCaMP) double transgene, allowing the expression of a genetically encoded calcium sensor in all neurons. We observed the rhythmic activation of calcium signaling in motor neurons located within the spinal cord (Fig. 4A, Movie S5). In wild-type larvae, the calcium signals exhibited an alternating pattern between the left and right sides of the body, whereas this coordinated pattern was disrupted in epha4a morphants (Fig. 4 A, B vs A’, B’, Movie S6). Additionally, we found that the activation frequency of motor neurons on the left and right sides was comparable in wild-type larvae, but significantly different in the absence of Epha4a (Fig. 4C-D).

Uncoordinated left-right activation of spinal cord neurons in the absence of Epha4a.

A, A’: Fluorescent images showing the dorsal-view of 24 hpf Tg(elavl3:GAL4; UAS:GCaMP) double transgenic larvae. The corresponding movies are shown in Movie S5 and S6. B, B’: Line charts showing the quantification of fluorescence changes of the region of interests (ROIs, circled in A, A’) in control and epha4a morphants. C: Scatter plot showing the distribution trend of the ratio of the calcium signal frequency between left and right in control (N=15 larvae, n=38 experiments) and epha4a morphants (N=15 larvae, n=41 experiments). D: Statistical graph of the ratio of the calcium signal frequency between left and right in control and epha4a morphants. Scale bars: 50 μm in panel (A, A’).

One well-established concept regarding left-right coordination involves the presence of central pattern generators (CPGs), which are regulated by interneuron circuitry within the spinal cord. We further examined the axon guidance of interneurons in epha4a mutants. First, we investigated the commissural trajectories of reticulospinal (RS) interneurons. In control larvae, the large Mauthner neurons, along with other RS neurons, were symmetrically positioned on each side of the midline and projected their axons to the contralateral sides (Fig. 5A). These bilaterally projected axons typically crossed at the midline and subsequently synapsed on motoneurons of the opposite sides, contributing to the generation of spinal cord neural circuits (Hale et al. 2016). However, in epha4a mutant larvae, we observed an abnormal pattern in axonal projections. Specifically, the mutant axons failed to traverse the midline and instead extended ipsilaterally (Fig. 5A). In addition, the distance between Mauthner neurons (rhombomere 4) and rhombomere 7 was significantly decreased and rhombomere 5 was scarcely visible in the mutant larvae (Fig. 5A-B). Furthermore, the sites of axon crossing between two Mauthner neurons tended to deviate to one side of the midline in the mutants (Fig. 5C-D’).

Disorganized neural patterning in epha4a mutants.

A: Confocal images showing reticulospinal neuronal axons in 48 hours post-fertilization (hpf) wild-type and epha4a mutant larvae visualized with anti-neurofilament antibody RMO44. Asterisks indicate the cross sites of Mauthner axons. The green arrow indicates the cell body of the r5/MiD2 neuron in a wild-type larva. The red arrow points to the ipsilaterally projected axon of r6/MiD3 in the mutant larva, which is normally projected to the other side in wild-type fish. B: Statistical chart showing the distance between r4 and r7 of 48 hpf wild-type and epha4a mutants. C: Scatter plot showing the distance (d) between the center line and the intersection site of Mauthner axons as indicated in panel C′. D: The ratio of the distance between the intersection site of Mauthner axons and bilateral axon bundles in 48 hpf wild-type (N=11 larvae) and epha4a mutants (N=12 larvae). The ratios were calculated as in panel D′. Scale bars: 20 µm in panel (A).

Cerebrospinal fluid-contacting neurons (CSF-cNs) represent a unique type of interneuron responsible for modulating the V0 and V2a interneurons, which are integral components of locomotor CPGs (Fidelin et al. 2015, Wu et al. 2021, Talpalar et al. 2013). In wild-type larvae, we observed that the ascending axons of these neurons projected either to the right or left side from the midline, as visualized using Tg(pkd2l1:GAL4;UAS:Kaede) double transgenic larvae (Fig. 6A). However, in the absence of Epha4a, the projection of these neurons exhibited notable disorganization, with numerous axons crossing the midline from one side of the trunk. This disorganized pattern was observed in both heterozygotic and homozygotic mutants (Fig 6A-B). Additionally, we employed an optogenetic approach to activate these CSF-cNs, utilizing the Tg(Gal4s1020t; UAS:ChR2) double transgene. Following optical stimulation, we observed robust tail oscillations as previously described (Wyart et al. 2009). In wild-type larvae, these tail oscillations exhibited a symmetrical left-right beating pattern(Fig. 6C, D, Movie S7). However, a striking disruption of this symmetry was observed in epha4a morphants, as they consistently beat towards one side of the trunk following optical stimulation(Fig. 6C’, D’ and E, Movie S8). Collectively, these findings demonstrate that the left-right coordination deficiencies observed in epha4a mutants arise from abnormal neural circuit formation, which consequently disrupts the integrity of the CPGs.

Aberrant swimming as a result of abnormal extension of CSF-cNs axons.

A: Fluorescent images showing the distribution of ascending axons of CSF-cNs marked by Tg(pkd2l1:GAL4,UAS:Kaede) in 2 dpf epha4a mutant larvae. Yellow line indicates the midline and the red arrows indicate aberrantly extended axons in epha4a+/- and epha4a-/- larvae. B: Bar graph showing the percentages of abnormal extension of CSF-cNs axons in 2 dpf wild-type and epha4a mutant; the numbers of larvae are indicated in each column. C, C’: Superimposed frames of tail oscillations in 5 dpf control and epha4a morphants. D, D’: A plot of body curvature angles in panel (C) and (C’). E: Percentages of tendentious swing in control (N=8 larvae, n=24 experiments) and epha4a morphants (N=11 larvae, n=33 experiments). The percentages were calculated by the ratio of tendentious tail oscillation during the first eight swings. Scale bars: 50 µm in panel (A).

Ephrin B3-EphA4 signaling regulates interneuron axon extension

To further explore the role of Epha4a during interneuron axon extension, we examined the expression of epha4a during early zebrafish embryonic development. Whole-mount in situ hybridization results showed that both epha4a and epha4b were abundantly expressed in the zebrafish spinal cord (Fig. S5A). We plotted the expression of these two genes using a published single-cell transcriptome data, which showed that epha4a and epha4b were both expressed in interneurons such as serotonergic interneurons as well as V0 and V2a interneurons, suggesting a role for epha4 in interneuron function (Fig. S5B) (Cavone et al. 2021). Notably, the expression of efnb3b, encoding the ligand of Epha4, was highly enriched in the midline floor plate cells (Fig. S5B).

Ephrins, through interacting with Eph receptors, play a critical role in repulsive axon guidance during neural development (Flanagan and Vanderhaeghen 1998, Egea and Klein 2007). We further analyzed axon guidance in efnb3b morphants. Similar to those of epha4a mutants, the efnb3b morphants also displayed axon guidance defects, as well as abnormal startle response and uncoordinated calcium activation (Fig. S6A-E). In addition, morphants larvae also displayed left-right oscillations defects after optogenetic stimulation (Fig. S6D-E).

Candidate variants in the EPHA4-related genes

Our zebrafish studies suggested that EPHA4 signaling is crucial for interneuron axon guidance, hence the formation of functional CPGs. Next, we further asked whether mutation of other components of the EPHA4 signaling can result in IS in humans. By searching for rare variants in EPHA4-related genes (Fig. S7) (Szklarczyk et al. 2019), we identified heterozygous de novo start-loss variant (c.1A>G, p.Met1?) in NGEF in a 17-year-old male (SCO2003P3332) (Fig. 7A-B, Table 1), whose scoliosis was diagnosed at 15 years of age with a main curve Cobb angle of 60°. NGEF encodes the neuronal guanine nucleotide exchange factor Ephexin that differentially affects the activity of GTPases RHOA, RAC1, and CDC42. The activation of Ephexin is triggered by ephrin through EPHA4 (Shamah et al. 2001).

IS patients with potential NGEF variants.

A: Protein structure of NGEF protein with the position of potential variants. B, C: Pedigrees and spinal radiographs of two probands with dominant gene variants. Sanger sequencing results are shown on the right.

Strikingly, in a European-ancestry IS cohort, we further identified a dominant missense variant (c.857G>A, p.Ala286Val) in NGEF in a quad family with three affected members (Fig. 7C, Table 1). The proband (II:2, TSRHC01) with onset of scoliosis at 14 years of age has a 58° main curve Cobb angle. This variant, which is located in the RhoGEF domain of Ephexin, is predicted to be highly deleterious (CADD=29.6).

Altogether, our results suggest that defects of the CPGs owing to abnormal EPHA4 signaling maybe one of the crucial factors responsible for IS.

Discussion

IS is a disease with diverse causes, and the underlying mechanisms can vary even among patients with similar scoliotic phenotypes. Previous genetic studies have highlighted the significance of the extracellular matrix (ECM) in maintaining the balance of axial bone and supporting soft tissues in the spine, thus playing a crucial role in IS development (Haller et al. 2016). For example, the top SNP associated with IS maps to LBX1, an essential molecule for ECM maintenance and bone homeostasis (Takahashi et al. 2011). Additionally, genetic loci in muscle development-related genes have also been associated with the onset of scoliosis, emphasizing the intricate interaction between bones and muscles (Ogura et al. 2015). However, it is worth noting that the variants in these genes explain only a small fraction of the overall heritability of IS. Consequently, it is imperative to establish connections between the extensive genetic findings and biological mechanisms that can elucidate the etiological landscape of IS.

In this study, we mapped 41 significant genome-wide loci to functional genes through positional mapping and functional mapping such as eQTL. Then we determined the enrichment in patient cohorts of rare variants in these genes, which may have greater impacts compared with common SNPs. This approach revealed the convergence of SNPs and rare variants in EPHA4 that are enriched in patients with IS. We also identified additional high-impact variants in NGEF, which is involved in the EPHA4 pathway.

Our studies using zebrafish have revealed that deficiency of Epha4 can lead to the development of scoliosis. Interestingly, we observed that even heterozygotic epha4a mutants displayed mild scoliosis (Fig. 2C), which is consistent with the occurrence of this condition in scoliosis patients. Further analysis involving behavior and imaging demonstrated that the absence of Epha4a resulted in defective left-right coordination. This coordination is crucially governed by Central Pattern Generators (CPGs), which generate rhythmic patterns of neural activity to coordinate limb movements on both sides of the body (Talpalar et al. 2013, Marder and Bucher 2001, Kiehn 2006). Previous studies have reported the involvement of ephrin and its receptors in axon guidance during the maturation of neural circuits, including CPGs (Iwasato et al. 2007, Borgius et al. 2014, Kullander et al. 2003, Andersson et al. 2012). In line with this, we observed disrupted axon guidance of the interneurons, which are integral components of CPGs, in epha4a mutants. This finding suggests that CPG malfunction is the primary factor underlying the scoliosis observed in these mutants. It is highly likely that the lack of coordinated left-right locomotion generates imbalanced mechanical forces on the spine, gradually leading to spinal curvature during later stages of development (Fig. 8).

The proposed mechanism of IS mediated by EPHA4 dysfunction.

In healthy individuals, EphrinB3-activated EPHA4 phosphorylates CDK5, leading to the phosphorylation of Ephexin, a protein encoded by NGEF. Phosphorylated Ephexin can regulate axon guidance through either activating RHOA or suppressing CDC42 and RAC1 signaling. These processes are critical to maintain the normal function of the CPG, the local neural network that provides coordinated bilateral muscle control. Impairment of the EPHA4 pathway and CPG may cause an imbalance of the motor drive from the spinal cord during development, thus causing the uncoordinated left/right swimming behavior in zebrafish larvae and the asymmetry of the bilateral muscular pull in a young child. Although the appearance is normal in early childhood, the dysfunction produces a scoliotic curve during the growth spurt.

Thus, our data provided a novel biological mechanism of IS, i.e., the impairment of neural patterning and CPG. Previous studies have provided clues on the role of CPGs in IS. Patients with adolescent IS showed asymmetric trunk movement during gait, as characterized by increased relative forward rotation of the right upper body in relation to the pelvis (Kramers-de Quervain et al. 2004, Nishida et al. 2017). An electromyography (EMG) study also showed asymmetric activation of paraspinal muscles between the convex and concave sides at the scoliosis curve apex (Shimode, Ryouji and Kozo 2003). In a child with a strong family history of IS, asymmetric hyperactivity was observed by EMG months before scoliosis was evident (Valentino et al. 1985). These left-right locomotor coordination abnormalities indicated the maldevelopment of CPGs as potential cause of IS. The CPG asymmetry may induce an imbalance in trunk muscle strength, resulting in asymmetric rib drooping. This leads to an abnormal vertebral rotation and then the onset of IS. Moreover, in a companion study, Wang et al. identified a number of rare variants in SLC6A9, which encodes glycine transporter 1 (GLYT1), in familial and sporadic adolescent IS cases. The slc6a9 mutant zebrafish also exhibited discoordination of spinal neural activities with pronounced lateral spinal curvature, recapitulating the human IS phenotype (data unpublished). Taken together, we propose that the dysfunction of CPGs would cause an imbalance in the motor drive from the spinal cord and the asymmetric transversospinalis muscle pull, eventually producing a scoliotic curve (Fig. 8).

In summary, we showed that both common and rare variants in the EPHA4 pathway contribute to the genetic architecture of IS. The dysfunction of the EPHA4 pathway causes IS through the impairment of neural patterning and CPGs.

Materials and Methods

Mapping of candidate genes utilizing previous association studies

Through a systematic literature review, we identified IS-associated SNPs reported in GWASs and meta-analyses of GWAS. The literature search was carried out using MEDLINE (via Pubmed.gov) and Web of Science (via Clarivate Analytics) and was limited to English-language articles published from January 1980 to October 2020. The following keywords were combined to perform the search: ‘idiopathic scoliosis’ AND ‘GWAS’ OR ‘SNP’ OR ‘single nucleotide polymorphism’ OR ‘variant’ (Table S4). The inclusion/exclusion criteria of abstracts are provided in Table S5. After screening the titles and abstracts, we obtained from the full-text articles the rsID, chromosome, and position for SNPs with a threshold for genome-wide significance of P < 5.0 × 10−8. For SNPs identified in multiple studies, we recorded the lowest P-value.

SNPs reported in previous studies were first pruned using Functional Mapping and Annotation (FUMA, v1.4.1, https://fuma.ctglab.nl/) (Watanabe et al. 2017). Significant SNPs were considered independent at r2 < 0.6. All known SNPs (available in the 1000 Genomes reference panel, https://www.internationalgenome.org/) were included for further gene mapping if they were in a linkage disequilibrium (LD) block (r2 ≥ 0.6) with a significant independent SNP. Three SNP-to-gene strategies were used:

For positional mapping of significant independent SNPs, we used annotations obtained from ANNOVAR (http://annovar.openbioinformatics.org). A 10-kb maximum distance was applied for intergenic SNPs.

For the eQTL mapping, we mapped significant independent SNPs and SNPs in an LD block to eQTLs acorss 44 GTEx tissue types (release V8) (Consortium 2020). SNP-gene pairs with a false discovery rate ≤ 0.05 were considered significant.

For chromatin interaction mapping, we overlapped the significant independent SNPs and SNPs in an LD block with one end of significantly interacting regions in all optional tissue/cell types. Genes were mapped if their promoter regions overlapped with another end of the significant interactions. The promoter region was defined as the region from −250 bp to +500 bp relative to the transcription start site.

Cohort description

The Peking Union Medical College Hospital (PUMCH) cohort: The PUMCH cohort comprised 411 unrelated Chinese patients with severe IS (Cobb angle ≥ 40°) who underwent spinal surgery in the PUMCH between October 2017 and March 2022 as part of the Deciphering disorders Involving Scoliosis and COmorbidities (DISCO) study (http://www.discostudy.org/). The clinical diagnosis was confirmed using standing full-spine radiographs, three-dimensional computed tomography, and magnetic resonance imaging. These patients did not show any congenital or neuromuscular defect at the time of recruitment. The control cohort consisted of 3,800 individuals without observable scoliosis and with exome sequencing or genome sequencing performed at PUMCH for clinical or research purposes. Individuals with vertebrae malformation or congenital developmental defects were excluded. A patient with a 2q35-36.2 deletion, including the EPHA4 gene, was also recruited through in-house gene matching of the DISCO study (Li et al. 2015). This patient was subsequently evaluated for scoliosis by examination and radiography.

The East Asian cohort: The East Asian cohort includes totally 6,449 IS patients and 158,068 controls from four independent datasets (Japanese dataset 1 (Kou et al. 2013, Takahashi et al. 2011): 1,261 cases, 15,019 controls; Japanese dataset 2 (Ogura et al. 2015): 878 cases, 21,334 controls; Japanese dataset 3 (Kou et al. 2019): 3,333 cases, 119,630 controls; Hong Kong dataset: 977 cases, 2,085 controls). The inclusion criteria for IS subjects was as same as our previous studies (Kou et al. 2019, Fan et al. 2012).

The Texas Scottish Rite Hospital for Children (TSRHC) cohort: We used a replication cohort of European-ancestry patients with IS from the TSRHC. Cases considered for inclusion in the study met criteria for a positive diagnosis of IS: lateral deviation from the midline greater than 15 degrees as measured by the Cobb angle method from standing spinal radiographs, axial rotation toward the side of the deviation and exclusion of relevant co-existing diagnoses.

Blood sample collection

In the PUMCH cohort, genomic DNA samples were extracted from peripheral blood leukocytes of each subject using a QIAamp DNA Blood Mini Kit (Qiagen, Hilden, Germany), according to the manufacturer’s protocol. Purified DNA was qualified by Nanodrop 2000 (Thermo Fisher Scientific, Waltham, MA, USA) and quantified by Qubit 3.0 using the dsDNA HS Assay Kit (Life Technologies, Carlsbad, CA, USA). DNA samples were stored at 4°C until used.

In the East Asian and TSRHC cohort, genomic DNA was extracted from peripheral blood or saliva using a standard protocol.

DNA sequencing and variant calling

In the PUMCH cohort, whole exome sequencing (WES) or whole genome sequencing (WGS) was performed on peripheral blood DNA from all individuals and available family members (Table S6). A SureSelect Human All Exon V6+UTR r2 core design (91 Mb, Agilent) was used for exon capture. The exomes were then sequenced on an Illumina HiSeq 4000 (Illumina, San Diego, CA, USA) according to the manufacturer’s instructions. For WGS, sequencing libraries were prepared using the KAPA Hyper Prep kit (KAPA Biosystems, Kusatsu, Japan) with an optimized manufacturer’s protocol. We performed multiplex sequencing using an Illumina HiSeq X-Ten sequencer (Illumina, San Diego, CA, USA). Variant calling and annotation were done in-house using the Peking Union Medical College Hospital Pipeline (PUMP) described previously (Zhao et al. 2021, Chen et al. 2021).

In the TSRHC cohort, subjects in 162 adolescent IS families were sequenced as part of the Gabriella Miller Kids First Pediatric Research Consortium (GMKF) at The HudsonAlpha Institute for Biotechnology (Huntsville, AL). In summary, DNA was normalized, sheared and then ligated to Illumina paired-end adaptors. The purified ligated DNA was amplified and exome sequencing was performed on the Illumina HiSeq X platform. The sample’s sequences were aligned into GRCh38 and genotypes were joint-called per each family by the GMKF’s Data Resource Center (DRC) at Children’s Hospital of Philadelphia following GATK best practices as detailed here (Mukhopadhyay et al. 2020). GMKF DRC’s alignment and joint genotyping pipelines are open source and made available to the public via GitHub (https://github.com/kids-first/kf-alignment-workflow and https://github.com/kids-first/kf-jointgenotyping-workflow).

Rare variant association analysis of the candidate genes

To determine the contribution of rare variants in GWAS candidate genes to IS, we analyzed the gene-based mutational burden for 156 candidate genes. Only ultra-rare variants with a gnomAD population-max allele frequency ≤ 0.01% and a cohort allele count ≤ 3 were analyzed. Each variant was assigned a weight, and the mutational burden of a given gene was defined as the maximum weight value among all ultra-rare variants carried by the individual. The SAIGE-GENE+ package was used to determine the weighted mutational burden test for each gene (Zhou et al. 2022).

Weighting criteria for the weighted burden analysis were developed according to the variant types and in silico results. The LoF variants (canonical splicing variants, nonsense variants, or variants that cause frameshift, stop-gain, or start-loss) were calculated together with the protein-altering variants, including missense variants and in-frame indels. Each variant was assigned a weight range from 0-1 and the mutational burden for a given gene was defined as the maximum weight value among all ultra-rare variants carried by the individual. The LoF variants annotated as ‘high confidence’ by the loss-of-function transcript effect estimator (LOFTEE) (Karczewski et al. 2020) were assigned a weight value of 1.0. The LoF variants annotated as ‘low confidence’ or unlabeled by LOFTEE, the non-canonical splicing variants with a SpliceAI score > 0.5, and the missense variants with a rare exome variant ensemble learner (REVEL) (Ioannidis et al. 2016) score of 0.8 were assigned a weight value of 0.8. The in-frame insertions/deletions (indels) with a Combined Annotation Dependent Depletion (CADD) (Kircher et al. 2014) score > 20 and the missense variants with a REVEL score > 0.6 and ≤ 0.8 were assigned a weight value of 0.6. The in-frame indels with a CADD score > 10 and ≤ 20 and the missense variants with a REVEL score > 0.4 and ≤ 0.6 were assigned a weight value of 0.4. The in-frame indels and the missense variants with a REVEL score > 0.2 and ≤ 0.4 were assigned a weight value of 0.2. The remaining missense variants were assigned a weight value of 0.

Sanger sequencing of familial participants

Sanger sequencing of familial participants was performed to determine the origin of variants in EPHA4 and NGEF. All LoF variants and protein-altering variants identified in familial participants were validated. Variant-encoding gene regions were amplified by PCR from genomic DNA obtained from probands, as well as from parents for trios, to determine the origin of the variants. The amplicons were purified using an Axygen AP-GX-50 kit (Corning, NY, USA) and sequenced by Sanger sequencing on an ABI 3730xl instrument (Thermo Fisher Scientific, Waltham, MA, USA).

Minigene assay

The splicing variant (EPHA4: c.1443+1G>C) was characterized by a minigene assay. Genomic DNA from the heterozygous patient was amplified by PCR using a high-fidelity DNA polymerase. Amplicons included exon 6, intron 6, exon 7, intron 7, and exon 8 of the EPHA4 gene. PCR products were cloned into the vector via the restriction sites BamHI and MluI for pCAS2, which is based on the pcDNA3.1 plasmid (Thermo Fisher Scientific, Waltham, MA, USA). Clones with wild-type or mutant genomic inserts were selected and verified by sequencing the cloned DNA fragments. The recombinant plasmids were transfected into 293T cells using LipofectamineTM 3000 reagent (Thermo Fisher Scientific, Waltham, MA, USA). For RT-qPCR, total RNA was isolated from the transfected cells using Trizol reagent (Thermo Fisher Scientific, Waltham, MA, USA and reverse transcription was performed using the GoScript™ Reverse Transcription System (Promega, Madison, MI, USA). PCR amplification was performed using the pCAS2-RT-F and pCAS2-RT-R primers, and the products were sequenced using pCAS2-RT-F.

Primers sequences for the minigene assay were as follows:

pCAS2-RT-F: 5′-CTGACCCTGCTGACCCTCCT-3′

pCAS2-RT-R: 5′-TTGCTGAGAAGGCGTGGTAGAG-3′

Nested PCR

The splicing variant (EPHA4: c.1318+10344A>G) was characterized by nested PCR. RNA was extracted from the whole blood of the patient using the TRIzon Reagent (CWBIO, Hangzhou, China) according to the manufacturer’s guidelines. The cDNA was synthesized using HiScript II 1st Strand cDNA Synthesis Kit with a gDNA wiper (Vazyme, Nanjing, China) according to the manufacturer’s guidelines. Nested PCR was performed as described previously (Yao and Tavis 2005). Nested PCR reactions in 50 μl, included 2 μl cDNA from the reverse transcription reaction as the template for the first round of PCR or 2 μl of first-round PCR product as the template for the second PCR, 1.5 μl 10 μM sense primer, 1.5 μl 10 μM anti-sense primer, 5 μl nucleotide mix (2 mM each dNTP), 5 μl 10× KOD Buffer, 1 μl 1 unit/μl Kod-Plus-Neo polymerase, 3 μl 25 mM MgSO4 and 31 μl ddH2O. The PCR program was (94°C for 2 min, 98°C for 10 s, 55°C for 30 s, 68°C for 60 s) × 20 cycles and then (94°C for 2 min, 98°C for 10 s, 57°C for 30 s, 68°C for 60 s) × 30 cycles.

Primers sequences:

For the first round:

F: 5′-GGCTCCTGTGTCAACAACTC-3′

R: 5′-GTTGGGATCTTCGTACGTAA-3′

For the second round:

F: 5′-AACTGCCTATGCAACGCTGG-3′

R: 5′-AGCTGCAATGAGAATTACC-3′

Western blots

For the missense variant in EPHA4 (c.2546G>A, p.Cys849Tyr), wild-type and mutant EPHA4 proteins were expressed from EPHA4-C1-pEGFP plasmids introduced into the 293T cells. Cells were cultured in six-well plates and transfected with DNA (2 mg/well) using Lipofectamine 3000 reagent (Thermo Fisher Scientific, Waltham, MA, USA). After 48 hours, cells were harvested, and protein extracts were prepared as described (Ding et al. 2017). The mutant proteins and wild-type proteins were fused to GFP. The expression of the two proteins was compared by western blots with an GFP antibody using the ECL detection system. The GFP antibody was purchased from Cell Signaling Technology (Cell Signaling, Danvers, MA, USA). The CDK5 antibody was purchased from Santa Cruz Biotechnology (Santa Cruz, CA, USA). A phospho-specific antibody against CDK5 phosphorylated at Tyr (GeneTex, Irvine, CA 92606 US) was purchased from GeneTex. Western blot experiments were repeated twice with similar results for the replicates.

Genotyping and imputation of GWAS

In the East Asian Japanese cohort, genotyping was performed by Illumina Human610 Genotyping BeadChip, Illumina HumanOmniExpressExome and HumanOmniExpress as our previous GWASs (Kou et al. 2019, Ogura et al. 2015, Takahashi et al. 2011, Kou et al. 2013). For quality control (QC), subjects with call rate < 0.98, high-degree relatedness with other subjects and outliers of East Asian ethnicity were excluded. For variants QC, the exclusion criteria of variants were as follows: call rate < 0.99, P-value for Hardy-Weinberg equilibrium < 1.0 × 10−6 and minor allele count < 10. The reference panel for imputation, namely JEWEL7K, was composed of 1000 Genomes Project phase 3 (v5) (Auton et al. 2015) and Japanese whole-genome sequence data (Das et al. 2016) with 3256 high-depth subjects (≥30 read counts) and 4216 low-depth subjects (≤ 15 read counts). Using EAGLE 2.4.1 (https://alkesgroup.broadinstitute.org/Eagle/) to determine the haplotypes, pre-phasing was conducted. Genotypes were imputed using Minimac4 (version1.0.0) (Das et al. 2016). After imputation, we excluded variants with minor allele frequency (MAF) < 0.005 and low imputation quality (R2 < 0.3).

In the East Asian Hong Kong cohort, samples were genotyped with Illumina Human Omni ZhongHua-8 Beadchips. Illumina Genome Studio v2.0 was used to convert raw data into PLINK format. The QC steps of samples and variants were described in a previous study (Marees et al. 2018). Genotype phasing and imputing were executed using SHAPEIT v2.r900 (Delaneau, Marchini and Zagury 2012) and IMPUTE2 (Marchini et al. 2007). The imputed data was filtered using the following parameters: INFO > 0.6, Certainty > 0.8, and MAF > 0.01. Association analysis was performed using PLINK v1.9 logistic regression with covariates: sex, age and top 20 principal components of the variance-standardized relationship matrix.

East Asian GWAS meta-analysis for IS

For the meta-analysis of the four datasets (three Japanese datasets and one Hong Kong dataset), an inverse-variance-based method was performed by METAL (version2011-03-25) (Willer, Li and Abecasis 2010). SNPs in three or more of the four cohorts were used in subsequent analyses.

Gene-based common variant analysis and eQTL analysis

SNPs in EPHA4 or within 20 kb flanking EPHA4 with a significant association with IS were retrieved. SNPs were matched to potential eQTLs according to the GTEx database (v8, https://gtexportal.org). Gene-based common variant analyses were performed using Multi-marker Analysis of GenoMic Annotation (MAGMA) (de Leeuw et al. 2015) and FUMA (Watanabe et al. 2017) using default settings with LD information from the 1000 Genomes Project East Asian population (1KGP EAS) as a reference. SNPs located within two kb upstream and one kb downstream from EPHA4 were included in the gene-based analysis.

Zebrafish strains, mutants, and morphants

Zebrafish Tuebingen (TU) strains were maintained at 28°C on a 14-hour/10-hour light/dark cycle. Embryos were raised at 28.5°C in E3 medium (5 mM NaCl, 0.17 mM KCl, 0.39 mM CaCl2, 0.67 mM MgSO4) following standard protocols. Zebrafish have two homologs of EPHA4, epha4a and epha4b. The CRISPR/Cas9 system was used to generate zebrafish epha4a and epha4b mutants. To increase efficiency, we injected Cas9 mRNA together with multiple single guide RNAs (sgRNAs) for each gene (Table S7). The sgRNA sequences for epha4a and epha4b are listed in Table S7. Morpholino sequences for epha4a and efnb3b knockdown analysis are also listed in Table S7.

Micro CT imaging

Adult zebrafish epha4a mutants or wild-type siblings were euthanized with tricaine methanesulfonate and fixed in 4% paraformaldehyde. Micro CT images were captured using a PerkinElmer Quantum GX2. Planar images acquired over 360° of rotation were reconstructed using QuantumGX. Three-dimensional renders of the skeleton were made with Analyze 12.0 software (AnalyzeDirect).

Behavioral recordings and analysis

For behavior analysis, individual zebrafish larvae were transferred into 24-well plate with fresh E3 medium at 8 dpf. Then, the plate was placed inside the Daniovision (Noldus) observation chamber for further behavior analysis. Video-tracking of swimming activity and further statistical analysis were performed using the EthoVision™ XT10 software. Behavioral data were shown as total swim distance (mm), average velocity (mm/s), average relative turn angle (degree), average absolute turn angle (degree) and average absolute angular velocity (degree/s) at a duration for 4 min.

For tactile stimulation, zebrafish larvae were placed in a concave slide, and tactile stimulation was performed with glass capillaries to the head or tail sides. Larval startle responses were recorded using a high-speed video camera (Mikrotron, EoSens Mini1) at 1000 fps. Automated analysis of larval movement was performed using the FLOTE software package (Jain et al. 2014).

For optogenetic studies, the Tg(Gal4s1020t; UAS:ChR2) transgenic embryos were injected with control, epha4a or efnb3b morpholinos at one cell stage. At 5 dpf, the head of injected larva was mounted in 1% low melting-point agarose (Sigma) and the body was exposed in glass-bottom dishes (WPI). Leica M165FC fluorescence microscope was used to irradiate with a blue laser at 488nm wavelength. Larval responses were recorded using a high-speed video camera (Mikrotron, EoSens Mini1) at 500 fps. Body curvature analysis of larval movement was performed using the FLOTE software package.

Whole-mount in situ hybridization and immunofluorescence

The primer sequences used to amplify epha4a, epha4b and rfng genes were listed in Table S7. Probe synthesis and whole-mount in situ hybridization were performed according to standard protocols. For reticulospinal neurons immunostaining, embryos were fixed in 2% trichloroacetic acid at 48 hpf for 3-4 h, washed twice in 0.5% Triton X-100 in PBS and blocked in 0.5% Triton X-100, 10% normal goat serum, 0.1% bovine serum albumin (Solarbio) in PBS for 1 h. The embryos were stained by monoclonal anti-neurofilament 160 antibody (Sigma-Aldrich) overnight at 4°C, then stained by goat anti-mouse Alexa Fluor 488 (Invitrogen) after washing.

Analysis of neuronal calcium signals activity

Control, epha4a or efnb3b morpholinos were injected into Tg(elavl3:GAL4; UAS:GCaMP) transgenic embryos at one cell stage. At 24 hpf, embryos were paralyzed with 0.5 mg/ml α-Bungarotoxin (AlomoneLabs), then mounted in 1% low melting-point agarose (Sigma) in glass-bottom dishes (WPI). Neuronal calcium signal images were collected with IXON-L-888 EMCCD camera equipped on Dragonfly 200 Spinning Disk Confocal Microscope using a 20×/0.55 objective within 1 min at a frame rate of 10 fps.

Images were analyzed with imageJ software. To quantify the change in fluorescence intensity, a region of interest (ROI) was defined, and the fluorescence intensity Ft of different frames was normalized to Ft/Fmin based on the minimum fluorescence intensity Fmin in all frames. To compare the left-right alternation pattern of neuronal calcium signals, the ratio of the left-side calcium signals frequency (fleft) to the right-side calcium signals frequency (fright) was logarithmically transformed (lg-transformed).

Quantification and statistical analysis

Statistical analyses were performed in SPSS (version 15.0). Unpaired Student’s t-test, Welch one-way ANOVA or two-way ANOVA followed by Turkey’s multiple comparison test were applied when appropriate. All experiments were replicated at least three times independently. P<0.05 was considered statistically significant.

Study approval

Approval for the study was obtained from the ethics committee at the Peking Union Medical College Hospital (JS-098, JS-2364), the medical ethics committee of the Keio University Hospital (No. 20080129), the ethical committee of RIKEN Yokohama Institute (No. H20-17(8)), and the Institutional Review Board of the University Texas Southwestern Medical Center (protocol STU 112010-150). Written informed consent was obtained from each participating individuals and families in the three cohorts. For the control group, the protocols were approved by the ethics committee at Peking Union Medical College Hospital. All zebrafish studies were approved by the Animal Care Committee of the Ocean University of China (Animal protocol number: OUC2012316).

eLife Assessment

The authors combined human genetic analysis with zebrafish experiments to produce evidence that alleles that impair the function of EPHA4 cause idiopathic scoliosis (IS), a common spinal deformity. The significance of the findings is important because the cellular and molecular mechanisms that contribute to IS remain poorly understood. The human genetic data are quite convincing whereas the zebrafish data, although supportive, are incomplete.

Joint Public Review:

Summary:

Idiopathic scoliosis (IS) is a common spinal deformity. Various studies have linked genes to IS, but underlying mechanisms are unclear such that we still lack understanding of the causes of IS. The current manuscript analyzes IS patient populations and identifies EPHA4 as a novel associated gene, finding three rare variants in EPHA4 from three patients (one disrupting splicing and two missense variants) as well as a large deletion (encompassing EPHA4) in a Waardenburg syndrome patient with scoliosis. EPHA4 is a member of the Eph receptor family. Drawing on data from zebrafish experiments, the authors argue that EPHA4 loss of function disrupts the central pattern generator (CPG) function necessary for motor coordination.

Strengths:

The main strength of this manuscript is the human genetic data, which provides convincing evidence linking EPHA4 variants to IS. The loss of function experiments in zebrafish strongly support the conclusion that EPHA4 variants that reduce function lead to IS.

Weaknesses:

The conclusion that disruption of CPG function causes spinal curves in the zebrafish model is not well supported. The authors' final model is that a disrupted CPG leads to asymmetric mechanical loading on the spine and, over time, the development of curves. This is a reasonable idea, but currently not strongly backed up by data in the manuscript. Potentially, the impaired larval movements simply coincide with, but do not cause, juvenile-onset scoliosis. Support for the authors' conclusion would require independent methods of disrupting CPG function and determining if this is accompanied by spine curvature. At a minimum, the language of the manuscript could be toned down, with the CPG defects put forward as a potential explanation for scoliosis in the discussion rather than as something this manuscript has "shown". An additional weakness of the manuscript is that the zebrafish genetic tools are not sufficiently validated to provide full confidence in the data and conclusions.

Author response:

The following is the authors’ response to the original reviews

Joint Public Review:

Idiopathic scoliosis (IS) is a common spinal deformity. Various studies have linked genes to IS, but underlying mechanisms are unclear such that we still lack understanding of the causes of IS. The current manuscript analyzes IS patient populations and identifies EPHA4 as a novel associated gene, finding three rare variants in EPHA4 from three patients (one disrupting splicing and two missense variants) as well as a large deletion (encompassing EPHA4) in a Waardenburg syndrome patient with scoliosis. EPHA4 is a member of the Eph receptor family. Drawing on data from zebrafish experiments, the authors argue that EPHA4 loss of function disrupts the central pattern generator (CPG) function necessary for motor coordination.

The main strength of this manuscript is the human genetic data, which provides convincing evidence linking EPHA4 variants to IS. The loss of function experiments in zebrafish strongly support the conclusion that EPHA4 variants that reduce function lead to IS.

The conclusion that disruption of CPG function causes spinal curves in the zebrafish model is not well supported. The authors' final model is that a disrupted CPG leads to asymmetric mechanical loading on the spine and, over time, the development of curves. This is a reasonable idea, but currently not strongly backed up by data in the manuscript. Potentially, the impaired larval movements simply coincide with, but do not cause, juvenile-onset scoliosis. Support for the authors' conclusion would require independent methods of disrupting CPG function and determining if this is accompanied by spine curvature. At a minimum, the language of the manuscript could be toned down, with the CPG defects put forward as a potential explanation for scoliosis in the discussion rather than as something this manuscript has "shown". An additional weakness of the manuscript is that the zebrafish genetic tools are not sufficiently validated to provide full confidence in the data and conclusions.

We highly appreciate the reviewer’s insightful comments and the acknowledgment of the main values of our study. We agree with the reviewer that further experiments are needed to fully establish the relationship between CPG and scoliosis. In response, we have revised the conclusion in the manuscript to better reflect this. Additionally, we conducted further analyses on the mutants to provide additional evidence supporting this concept.

Reviewer #1 (Recommendations for the authors):

Epha4a mutant zebrafish exhibited mild spinal curves, mostly laterally and in the tail. This was 75% of homozyous mutants but also, surprisingly, about 20% of heterozygotes. epha4b mutants also developed some mild scoliosis. If the two zebrafish paralogs can compensate for each other (partial redundancy), we might expect more severe scoliosis in double mutants. Did the authors generate and analyze double mutants? I believe it would be very useful for this study to report the zebrafish phenotype of loss of both paralogs together.

We appreciate the reviewer’s insightful comment regarding the potential value of reporting the phenotype of eph4a/eph4b double mutants. While we fully agree that this analysis would be valuable, our attempts to generate double mutants have been unsuccessful. These two genes are closely linked on the chromosome, with less than 100 kb separating them, which makes it challenging to generate double mutants through standard genetic crossing. Establishing a double mutant line would require more than a year due to the technical constraints of the process. Although we are unable to address this question directly at this time, we hypothesize that eph4a/eph4b double mutants may exhibit a higher likelihood of body axis abnormalities based on the phenotypes observed in single mutants and the known functions of these genes.

We hope this perspective will provide some useful context despite the limitations.

In Figure 1F, a pCDK5 western blot is performed as a readout of EPH4A signaling after either WT or C849Y mutant EPH4A is transfected into HEK 293T cells. It would be useful to mention in the text, or at least the figure legend, how this experiment was performed/where the protein samples came from. It is included in the methods, but in the main text, it simply says "we conducted western blotting" without mentioning whether the protein samples were from cell lines, patients, or another source.

Sorry for our ignorance. A detailed description of the western blotting conduction was supplemented at both “results” part (page 8, line 187-190) and the Figure 1 legend.

Was the relative turn angle biased to the left or right side of the fish? (i.e. is a positive angle a rightward or leftward turn?)

We are sorry for our unclear description. In Figure 3D, positive angle means turning left, while negative angle means turning right. In wild-type larvae, the average turning angle over a 4-minute period is approximately 0, whereas in mutants, this value deviates from 0, indicating a directional preference (positive for leftward and negative for rightward turns) in swimming behavior during the recording period. We have also made the necessary supplementation in the text and figure legend.

In Figure 4, morpholinos rather than mutants are used, but it is not clear why. Has it been established that the MO used disrupts gene function specifically? Can the effect of the MO be rescued by expressing a wild-type mRNA of Epha4a? Does MO knockdown induce spinal curves if fish are raised? Indeed, this could be a way to determine whether the spinal curves are caused by early events in development (when MOs are active).

Thanks for the comments. The efficacy of relevant MOs has been well-documented in numerous previous studies (Addison et al., 2018; Cavodeassi et al., 2013; Letelier et al., 2018; Royet et al., 2017). Following this reviewer’s suggestion, we have raised the epha4a morphants into adults, while no scoliosis were observed, suggesting that the spinal curvature formation may be induced by long-term defects in the absence of Epha4a. Additionally, we reconfirmed the abnormal motor neuron activation frequency phenotype in the mutants background. The corresponding data have replaced the original Figure 4 in the manuscript.

References

(1) Addison, M., Xu, Q., Cayuso, J., and Wilkinson, D.G. (2018). Cell Identity Switching Regulated by Retinoic Acid Signaling Maintains Homogeneous Segments in the Hindbrain. Dev Cell 45, 606-620 e603.

(2) Cavodeassi, F., Ivanovitch, K., and Wilson, S.W. (2013). Eph/Ephrin signalling maintains eye field segregation from adjacent neural plate territories during forebrain morphogenesis. Development 140, 4193-4202.

(3) Letelier, J., Terriente, J., Belzunce, I., Voltes, A., Undurraga, C.A., Polvillo, R., Devos, L., Tena, J.J., Maeso, I., Retaux, S., et al. (2018). Evolutionary emergence of the rac3b/rfng/sgca regulatory cluster refined mechanisms for hindbrain boundaries formation. Proc Natl Acad Sci U S A 115, E3731-E3740.

(4) Royet, A., Broutier, L., Coissieux, M.M., Malleval, C., Gadot, N., Maillet, D., Gratadou-Hupon, L., Bernet, A., Nony, P., Treilleux, I., et al. (2017). Ephrin-B3 supports glioblastoma growth by inhibiting apoptosis induced by the dependence receptor EphA4. Oncotarget 8, 23750-23759.

Reviewer #2 (Recommendations for the authors):

Supplementary Table 3 is missing.

Sorry for any inconvenience caused to the reviewers. Due to the size of the supplementary Table 3, we have separately uploaded an Excel file as supplementary materials. We have also double-checked during the resubmission process of the revised manuscript. Thanks for your thorough review.

The authors report only a single mutant allele for zebrafish epha4a and epha4b. Additionally, they provide no information about how many generations each allele has been outcrossed. The authors should provide some type of validation that the phenotypes they describe result from loss of function of the targeted gene and not from an off-targeting event.

Thanks for the comments. For epha4a and epha4b mutants, each homozygous mutant was initially derived from the self-crossing of first filial generation heterozygotes, and subsequent homozygous generations were maintained for fewer than three rounds of in-crossing. Interestingly, we observed a reduction in the incidence of scoliosis across successive generations. This trend may be attributed to potential genetic compensation mechanisms, which could mitigate the phenotypic severity over time. To address concerns about possible off-target effects, we synthesized and injected epha4a mRNA to test for phenotypic rescue. Our data show that epha4a mRNA injection partially restored swimming coordination in the mutants (Fig. S5). Moreover, similar motor coordination defects have been reported in Epha4-deficient mice, as documented in previous studies (Kullander et al., 2003; Borgius et al., 2014). These findings collectively strengthen the hypothesis that Epha4a plays a critical role in regulating motor coordination.

References

(1) Borgius, L., Nishimaru, H., Caldeira, V., Kunugise, Y., Low, P., Reig, R., Itohara, S., Iwasato, T., and Kiehn, O. (2014). Spinal glutamatergic neurons defined by EphA4 signaling are essential components of normal locomotor circuits. J Neurosci 34, 3841-3853.

(2) Kullander, K., Butt, S.J., Lebret, J.M., Lundfald, L., Restrepo, C.E., Rydstrom, A., Klein, R., and Kiehn, O. (2003). Role of EphA4 and EphrinB3 in local neuronal circuits that control walking. Science 299, 1889-1892.

The authors need to provide allele designations for the mutant alleles following accepted nomenclature guidelines.

Thank you for your careful review! We have reviewed and made revisions to the genes and mutation symbols throughout the entire text.

The three antisense morpholino oligonucleotides need to be validated for efficacy and specificity.

Thanks for the comments. The morpholinos were extensively used and validated in previous studies, and the efficacy of these morpholinos has been thoroughly validated in multiple studies (Addison et al., 2018; Cavodeassi et al., 2013; Letelier et al., 2018; Royet et al., 2017). Furthermore, we also performed swimming behavior analysis in the mutant background, which showed similar results as the morphants. Moreover, we also performed rescue experiments to confirm the specificity of the mutants (Fig. S5). Finally, we reconfirmed the abnormal calcium signaling in the mutants (Fig. 4), which further support our previous knockdown results.

References

(1) Addison, M., Xu, Q., Cayuso, J., and Wilkinson, D.G. (2018). Cell Identity Switching Regulated by Retinoic Acid Signaling Maintains Homogeneous Segments in the Hindbrain. Dev Cell 45, 606-620 e603.

(2) Cavodeassi, F., Ivanovitch, K., and Wilson, S.W. (2013). Eph/Ephrin signalling maintains eye field segregation from adjacent neural plate territories during forebrain morphogenesis. Development 140, 4193-4202.

(3) Letelier, J., Terriente, J., Belzunce, I., Voltes, A., Undurraga, C.A., Polvillo, R., Devos, L., Tena, J.J., Maeso, I., Retaux, S., et al. (2018). Evolutionary emergence of the rac3b/rfng/sgca regulatory cluster refined mechanisms for hindbrain boundaries formation. Proc Natl Acad Sci U S A 115, E3731-E3740.

(4) Royet, A., Broutier, L., Coissieux, M.M., Malleval, C., Gadot, N., Maillet, D., Gratadou-Hupon, L., Bernet, A., Nony, P., Treilleux, I., et al. (2017). Ephrin-B3 supports glioblastoma growth by inhibiting apoptosis induced by the dependence receptor EphA4. Oncotarget 8, 23750-23759.

Line 229. "While in consistent with previous reports, the hindbrain rhombomeric boundaries were found to be defective....". This sentence is not clear. Please describe how it is "inconsistent".

Thanks for the comments and sorry for the unclear description, we have described this more clearly in our revised manuscript (page 9, line 229-230).

Animals frequently are described as "heterozygous mutants" or "mutants". Please make clear that the latter are homozygous mutant animals.

Thanks for the comments. In the manuscript, all references to mutants specifically indicate homozygous mutants. Heterozygous mutants are explicitly identified as such.

The chromatin interaction portion of the Methods does not include any information on how these experiments were conducted or where the data were obtained. This information needs to be provided.

Thanks for your advice. The detailed information of chromatin interaction mapping has been provided in “Methods and Materials” (page 18-19, line 450-455). Information about the interacting regions was derived from Hi-C datasets of 21 tissues and cell types provided by GSE87112. The significance of interactions for Hi-C datasets was computed by Fit-Hi-C, with an FDR ≤ 10-6 considered significant.

The authors present single-cell RNA-seq data in Supplementary Figure 5 for which they cite Cavone et al, 2021. This seems like an odd database to use. Can the authors provide an explanation for choosing it? In any case, the citation should also be made in the Supplementary Figure 5 legend.

Thank you for your rigorous comment, we have cited this literature in the proper place of the revised manuscript. Cavone et al. used the her4.3:GFP line to label ependymo-radial glia (ERG) progenitor cells and performed single-cell RNA-seq on FACS-isolated fluorescent cells. The isolated cells included not only ERG progenitors but also undifferentiated and differentiated neurons and oligodendrocytes. The authors attributed this to the relative stability of the GFP protein, which remained in the progeny of GFP-expressing her4.3+ ERG progenitor cells, thus effectively acting as a short-term cell lineage tracer. Indeed, clustering analysis of this data successfully identifies neural progenitors and other neural clusters. Therefore, we consider that this scRNA-seq data encompasses a comprehensive range of neural cell types and is suitable for analyzing the expression of genes of interest. Furthermore, we downloaded and analyzed the scRNA-seq data of the zebrafish nervous system reported by Scott et al. in 2021 (Fig. S7B) (Scott et al., 2021). Despite differences in the developmental stages of the larvae analyzed (Cavone et al. examined larvae at 4 dpf, whereas Scott et al. analyzed larvae at 24, 36, and 48 hpf), our findings are consistent. Specifically, epha4a and epha4b are expressed in interneurons, whereas efnb3a and efnb3b are enriched in floor plate cells.

References

(1) Scott, K., O'Rourke, R., Winkler, C.C., Kearns, C.A., and Appel, B. (2021). Temporal single-cell transcriptomes of zebrafish spinal cord pMN progenitors reveal distinct neuronal and glial progenitor populations. Dev Biol 479, 37-50.

In Figure Legend 1, "expressed from the EPHA4-mutant plasmid" is not an accurate description of the experiment.

Sorry for the previous inaccurate description. The description has been revised to accurately reflect the experiment. “Western blot analysis of EPHA4-c.2546G>A variant showing the protein expression levels of EPHA4 and CDK5 and the amount of phosphorylated CDK5 (pCDK5) in HEK293T cells transfected with EPHA4-mutant or EPHA4-WT plasmid”.

Figure 3 panels J and K need more explanation. I don't understand what the different colors represent nor do I understand what are wild type and what are mutant data.

Thank you for your valuable feedback. We apologize for the lack of clarity in the original figure legend. To address this, we have revised the legend of Figure 3 to provide a more detailed explanation. In panels J and K, each color-coded curve represents the response of an individual larva from an independent experimental trial to the stimulus. Specifically, panel J depicts the response data for the wild-type larvae, whereas panel K presents the response data for the homozygous epha4a mutants.

Please provide the genotypes for the images in Figure 5A.

Thanks for the comments and we are sorry for our unclear description, we have described this more clearly in the Figure 5.

Figure legend 6B should also note the heterozygote data with the wild type and homozygous mutant data.

Thanks for the comments, the data are now included in Figure 6B.

Epha4 and Efnb3 have well-established roles in axon guidance. Although this is noted in the Discussion, I think a more extensive description of prior findings would be helpful.

Thanks for your valuable feedback. A more detailed description of the roles of Epha4 and Efnb3 in axon guidance was provided in the “Discussion” (page 16, line 388-396).

The main conclusion of this manuscript is that EPHA4 variants cause IS by disrupting central pattern generator function. I think this is misleading. I think that the more valid conclusion is that EPHA4 loss of function causes axon pathfinding defects that impair locomotion by disrupting CPG activity, thereby leading to IS. I urge the authors to consider this more nuanced interpretation.

Thank you for your insightful comments. We appreciate your suggestion to refine our main conclusion. We agree that the proposed revision more accurately reflects our findings and will revise the manuscript accordingly to state that “EPHA4 loss of function causes axon pathfinding defects, which impair locomotion by disrupting central pattern generator activity, potentially leading to IS.”

Version 2

Abstract

Abstract

Idiopathic scoliosis (IS) is the most common form of spinal deformity with unclear pathogenesis. In this study, we firstly reanalyzed the loci associated with IS, drawing upon previous studies. Subsequently, we mapped these loci to candidate genes using either location-based or function-based strategies. To further substantiate our findings, we verified the enrichment of variants within these candidate genes across several large IS cohorts encompassing Chinese, East Asian, and European populations. Consequently, we identified variants in the EPHA4 gene as compelling candidates for IS. To confirm their pathogenicity, we generated zebrafish mutants of epha4a. Remarkably, the zebrafish epha4a mutants exhibited pronounced scoliosis during later stages of development, effectively recapitulating the IS phenotype. We observed that the epha4a mutants displayed defects in left-right coordination during locomotion, which arose from disorganized neural activation in these mutants. Our subsequent experiments indicated that the disruption of the central pattern generator (CPG) network, characterized by abnormal axon guidance of spinal cord interneurons, contributed to the disorganization observed in the mutants. Moreover, when knocked down efnb3b, the ligand for Epha4a, we observed similar CPG defects and disrupted left-right locomotion. These findings strongly suggested that ephrin B3-Epha4 signaling is vital for the proper functioning of CPGs, and defects in this pathway could lead to scoliosis in zebrafish. Furthermore, we identified two cases of IS in NGEF, a downstream molecule in the EPHA4 pathway. Collectively, our data provide compelling evidence that neural patterning impairments and disruptions in CPGs may underlie the pathogenesis of IS.

Introduction

Idiopathic scoliosis (IS) is the most common form of spinal deformity, affecting about 2.5% of the global population (Hresko, 2013; Luk et al., 2010). IS may have long-term physical and mental health consequences, such as cosmetic deformity, cardiopulmonary impairment, and even disability (Weinstein et al., 2003). All these consequences can severely reduce the quality of life. Early intervention with conservative treatments, such as braces, can control scoliosis progression and reduce the need for surgical intervention (Weinstein et al., 2013). However, IS often remains undiagnosed until malformation is evident, emphasizing the importance of risk-prediction measurements in medical management.

Genetic factors are thought to play a significant role in the development of IS, while only a few genes have been associated with the condition to date (Cheng et al., 2015; Miller, 2007). Several single-nucleotide polymorphisms (SNPs) associated with susceptibility to IS have been identified through genome-wide association studies (GWASs), including SNPs linked to genes such as LBX1, GPR126, and BNC2. Notably, knockdown or overexpression of zebrafish homologs of these IS-associated genes have yielded body axis defects (Guo et al., 2016; Kou et al., 2019; Kou et al., 2013; Ogura et al., 2015; Zhu et al., 2015). In addition to common SNPs, rare variants with larger effect sizes or causing rare Mendelian disease traits also contribute to IS. Rare variants in FBN1, FBN2, and other extracellular matrix genes are associated with severe IS (Buchan et al., 2014; Haller et al., 2016). Linkage analysis of familial IS suggests that it follows autosomal dominant (AD), X-linked dominant, and multifactorial patterns of inheritance (Miller et al., 2005). Variants in CHD7 and AKAP2 are also implicated in the pathogenesis of Mendelian forms of IS (Gao et al., 2007; Liebeskind et al., 2016). Finally, centriolar protein POC5 and planar cell polarity protein Vang-like protein 1 (VANGL1) were also shown to be associated with IS. Notably, all these genes participate in a wide range of biological processes, the mechanisms underlying IS progression still require further investigation. Consequently, there is currently no consensus on the etiology of IS (Tang et al., 2021).

In this study, we developed a novel pipeline that combines SNP-to-gene mapping and rare variant association analysis. Using this approach, we analyzed a large Chinese population with IS and further validated our findings in East Asian and European populations. Through these analyses, we identified EPHA4 as a novel gene associated with IS. To confirm the pathogenicity of this gene, we used a zebrafish model and found that impaired EPHA4 pathway components and resulting defects in central pattern generators (CPGs) were associated with asymmetric spinal locomotion, which may serve as a contributory factor in the development of IS. Additionally, we searched for candidate genes related to EPHA4 signaling pathways and identified mutations in the NGEF among IS patients. Overall, our data suggest that the impairment of the EPHA4 pathway and central pattern generators plays a previously unknown role in the development and progression of IS.

Results

Enrichment analysis of rare variants in IS candidate genes

To identify candidate genes associated with IS, we conducted a comprehensive literature review of SNPs that are linked to IS from 14 published GWASs. This led us to obtain forty-one SNPs that showed genome-wide significance (P < 5×10−8) as detailed in Table S1. We then employed positional mapping, expression quantitative trait locus (eQTL) mapping, or chromatin interaction mapping to identify 156 candidate genes that could be potentially linked to these SNPs. Further analysis of rare variants for the 156 candidate genes using exome data from 411 Chinese probands and 3,800 unrelated Chinese controls identified EPHA4 as the only significant gene (P=0.045, OR=4.09) (Table S2). EPHA4 is a member of the Eph family of receptor tyrosine kinases which play a vital role in the development of the nervous system (Kania and Klein, 2016). We identified three rare variants in EPHA4 from three IS patients, including one splicing-donor variant and two missense variants (Table S2). The flowchart of the entire pipeline is shown in Fig. S1.

Inheritance pattern and functional analyses of variants in EPHA4

We characterized the inheritance pattern of variants in EPHA4 using either trio sequencing data or Sanger sequencing for the parents. None of the parents being tested had scoliosis based on a clinical screening test. Of the three rare variants in EPHA4, two (NM_004438.3: c.1443+1G>C and c.2546G>A [p.Cys849Tyr]) arose de novo (Fig. 1A and Table 1). The heterozygous splicing variant (c.1443+1G>C) was identified in a female IS patient (SCO2003P0846) (Fig. 1B). Scoliosis in this patient was diagnosed in her 10s, with a main curve Cobb angle of 60°. The in vitro minigene assay showed that the c.1443+1G>C variant introduced a new splicing site, resulting in a 36-bp in-frame deletion in exon 6 (Fig. 1G, 1I, S2A). The EPHA4 heterozygous missense variant (c.2546G>A, p.Cys849Tyr), identified in a female IS patient (SCO2003P2146) (Fig. 1C), is located in the protein tyrosine kinase domain of EPHA4 protein (Fig. 1A). The onset of scoliosis in this patient was in her 10s with a 70° main curve Cobb angle. Given the critical role of EPHA4 in phosphorylating CDK5, which in turn activates downstream signaling pathways (Fu et al., 2007), a western blot analysis of EPHA4-c.2546G>A variant was conducted to investigate the protein expression levels of EPHA4 and CDK5 and the amount of phosphorylated CDK5 (pCDK5) in HEK293T cells transfected with EPHA4-mutant or EPHA4-WT plasmid. Our results revealed that the missense variant resulted in a decreased phosphorylation level of CDK5 (p=0.015), suggesting a partial loss-of-function (LoF) of EPHA4 (Fig. 1F).

Clinical and genetic information on IS patients and functional effect of EPHA4 variants.

A: Locations of three EPHA4 single nucleotide variants relative to the protein domains (top panel) and exons 1-18 (bottom panel). B-E: Pedigrees and spinal radiographs of four probands with dominant gene variants. Sanger sequencing confirmed the variants. The arrows indicate the probands. The term +/+ denotes the wild-type, and cDNA change/+ denotes the heterozygous variant. F: Western blot analysis of EPHA4-c.2546G>A variant showing the protein expression levels of EPHA4 and CDK5 and the amount of phosphorylated CDK5 (pCDK5) in HEK293T cells transfected with EPHA4-mutant or EPHA4-WT plasmid. WT: wild type. G: Schematic representation of the effect of the EPHA4-c.1443+1G>C mutation on the splicing process. This variant induced a new splicing site (red box). The yellow box indicates the splicing acceptor. H: Schematic representation of the effect of the EPHA4-c.1318+10344A>G mutation on the splicing process. This variant induced a new splicing site (red box). The yellow box indicates the splicing acceptor. I: The minigene assay result showed that the c.1443+1G>C variant introduced a new splicing site, resulting in a 36-bp in-frame deletion in exon 6. J: The nested PCR showed that the c.1318+10344A>G variant induced exon 5 skipping, resulting in a 339-bp in-frame deletion. K: NCBI RefSeq genes included in 2q35-q36.2 from UCSC Genome Browser. EPHA4 is shown by the red oval.

We next investigated de novo non-coding variants of EPHA4 via whole genome sequencing (WGS) in 116 trio families with IS. A de novo heterozygous EPHA4 intronic variant (c.1318+10344A>G) was identified in a female patient (SCO2003P2080) in an age range of 15-20 (Fig. 1D and Table 1). She developed scoliosis in her 10s with a 70° main curve Cobb angle. This variant is predicted to affect the branch point of the fifth intron of EPHA4 (Fig. S2B). We performed nested PCR to show that this variant induced exon 5 skipping, resulting in a 339-bp in-frame deletion (Fig. 1H, J).

Next, we employed an in-house gene matching approach under the framework of DISCO study, which identified a patient (SCO2003P3202) in an age range of 5-10, who had been previously diagnosed with Waardenburg syndrome caused by a 4.46 Mb de novo deletion at 2q35-q36.2 (Li et al., 2015) (Table 1). Intriguingly, the patient also presented mild scoliosis (Fig. 1E). The deletion included the entire PAX3 gene, which was responsible for the Waardenburg syndrome phenotype, and 36 neighboring genes, including EPHA4 (Fig. 1K). As scoliosis is not typically associated with Waardenburg syndrome caused solely by PAX3 pathogenic variants (Tassabehji et al., 1993), we hypothesize that the deletion of EPHA4 may be responsible for the IS phenotype in this patient. Notably, the patient IDs used in this study are unique to our research group and will not be recognizable to anyone outside the research team, including hospital staff and the patients themselves.

Notably, the GWAS signal which we mapped to EPHA4 (rs13398147) (Zhu et al., 2015) represents a significant eQTL in esophagus and colon tissues, with the T allele associated with decreased expression of EPHA4. In our East Asian GWAS cohort of 6,449 adolescent IS patients and 158,068 controls, we identified another two eQTLs in EPHA4 associated with decreased expression of EPHA4 in brain tissue (Table S3). In the same GWAS cohort, common SNPs in EPHA4, after aggregation, also showed significant enrichment (P=0.023) in IS patients versus controls. Taken together, the convergence between rare and common variants of EPHA4 that lead to LoF or hypomorphic effects highlights the pivotal role of EPHA4 in the pathogenesis of IS.

IS-like phenotypes in zebrafish epha4 mutants

To investigate the role of EPHA4 in scoliosis, we utilized zebrafish as a model system due to its versatile nature in modeling adolescent IS (Bagnat and Gray, 2020; Boswell and Ciruna, 2017; Grimes et al., 2016; Xie et al., 2022). Zebrafish have two homologs of EPHA4, epha4a and epha4b. Using CRISPR-Cas9, we established a stable epha4a zebrafish mutant line with a 63-bp deletion in exon 3, which introduced a stop codon and resulted in a truncated protein (Fig. 2A-B). The homozygous epha4a mutant larvae had no apparent defects in either notochord or body axis development (Fig. S3A). Consistent with previous reports, the hindbrain rhombomeric boundaries were found to be defective in both the epha4a homozygous mutants and morphants (Fig. S3B) (Cayuso et al., 2019; Cooke et al., 2005). Interestingly, more than 75% adult mutants showed mild scoliosis (88 of 116), and some mutants exhibited severe scoliotic phenotype (4 of 116) (Fig. 2C-D, Movie S1&S2). Remarkably, some heterozygous adult mutants also developed mild scoliosis (14 of 95), whereas none of the wild-type fish showed any signs of scoliosis (0 of 76) (Fig. 2C-D). Similarly, we further generated the epha4b mutants with a 25-bp deletion in exon 3, which resulted in a frameshift mutation (Fig. S4A). The epha4b zebrafish mutants also developed mild scoliosis (28 of 43) (Fig. S4B-C). Intriguingly, both epha4a and epha4b mutants exhibited early onset scoliosis starting from around 20 days post fertilization (Fig. S4D), a stage similar to that of IS patients. Collectively, these data suggest that mutations in Epha4 proteins are linked to the scoliotic phenotype in zebrafish.

Scoliosis in zebrafish epha4a mutants.

A: Diagram of the protein domains, genomic structures, and sequences of wild-type and corresponding epha4a mutants. Red arrows indicate mutation sites. Blue boxes indicate open reading frames. Underlined sequences indicate the protospacer adjacent motif (PAM) region, and red fonts indicate Cas9 binding sites. LBD: ligand binding domain; FN3: fibronectin type 3 domain; TMD: transmembrane domain; PTKc: catalytic domain of the protein tyrosine kinases; SAM: sterile alpha motif. B: Sanger sequencing results confirmed the deletion of the target region in epha4a mutant transcripts. C: Bar graph showing the percentages of adult zebrafish with normal, mild, or severe body axis defects in wild-type (n=76), epha4a heterozygote (n=95), or epha4a homozygous (n=116). D: Representative images of wild-type and homozygous epha4a mutants. Micro CT images are shown on the right. Lateral and dorsal views are shown. Scale bar: 1 cm.

Abnormal left-right swimming pattern in the absence of Epha4a

While the epha4a mutants seemed to be grossly normal during the larvae stages, they developed spinal curvature gradually during later development. We decided to investigate whether these mutants exhibited any abnormalities in their behavior at the larvae stages using EthoVision XT software. By monitoring the swimming behavior of 8 days post-fertilization (dpf) zebrafish larvae, we found the motion distance and swimming velocity were significantly decreased in epha4a mutants (Fig. 3A-B). In addition, we observed a remarkable difference in the relative turning angle and angular velocity between these two groups. The wild-type group changed their swimming direction randomly, showing a relative angle around 0 degrees (Fig. 3C-D). In contrast, the epha4a mutants favored turning to one side of their directions (a positive angle reflects a leftward turn) (Fig. 3C-D). Moreover, the absolute turn angles and turning speed (angular velocity) were significantly higher in these mutants (Fig. 3E-F). To rule out potential off-target effects, we injected epha4a mRNA into epha4a mutants, which significantly restored the swimming coordination defects (Fig. S5).

Abnormal left-right swimming pattern in the absence of Epha4a.

A: Dot plots showing the swimming distance of each 8 dpf larva at a duration of 4 min (N=54 for wild-type and N=60 for epha4a mutants). B: Dot plots showing the swimming velocity of wild-type and mutant larvae as indicated. C: Diagram showing the turning angle (θ) of the larvae during swimming. D: Scatter plot showing the relative turning angle of wild-type and mutant larvae. The relative angles were calculated by the sum of turning angles during fish swimming with left (positive) or right (negative) turns. The epha4a mutants favored turning to one side of their directions compared with those of wild-type larvae. E: Dot plots showing the average absolute turning angle of wild-type and mutant larvae as indicated. F: Dot plots showing the average absolute angular velocity of wild-type and mutant larvae as indicated. G: Bar graph showing the percentages of 5 dpf zebrafish larvae with different reactions after tactile stimulation. N=10 for each group; the numbers of tactile stimulations are indicated in each column. H: Representative images of the total body curvature measurements in zebrafish larvae, with values as the sum of α and β angles shown in the figure. I: Representative time-series images of 5 dpf wild-type and epha4a mutant zebrafish larvae after tactile stimulation to the head. Each panel represents the points of maximal body curvature for the C-bend and counterbend after the tactile startle response. J, K: A plot of body curvature angles as measured in panel (H) during swimming in response to tactile head stimulation in 5 dpf wild-type (Panel J, N=5 larvae, n=10 stimuli) and epha4a mutants (Panel K, N=5 larvae, n=7 stimuli). Each colored curve represents an independent experiment showing the response of a single larva to a stimulus. The positive angle means turning right. L: The maximum curvature angles during the first C-bend and counterbend after tactile stimulation in wild-type and epha4a mutant larvae.

Next, we compared the swimming behavior after startle response between wild-type and epha4a mutants. We used a needle to touch the head or the tail of 5 days post-fertilization (dpf) zebrafish larvae to stimulate swimming behavior. In sibling controls, the larvae responded to the tactile stimulation and swam away quickly (Movie S3). Conversely, the epha4a mutants failed to respond to the initial stimulation, and the swimming pattern was defective with an abnormal bending pattern (37 of 62 from tail-stimulation group and 34 of 77 from head-stimulation group) (Fig. 3G, Movie S4). We further performed more comprehensive analysis by high-speed video microscopy. After tactile stimulation, wild-type larvae displayed a high-speed C-bend turn followed by a weaker counterbend turn after several milliseconds (Fig. 3H-I). This rhythmic left/right swimming pattern ensures that the fish swim away from the frightening stimulus. The turning angles of the control larvae after stimulation changed with a sinusoidal wave pattern (Fig. 3J). In contrast, this pattern was dramatically different in epha4a mutants (Fig. 3I,K). Of note, although the C-bend turning angles were similar between epha4a mutants and control siblings, the turning angles of the counterbend decreased significantly (Fig. 3L), implying a left/right coordination defect. Taken together, both regular swimming and tactile stimulation analyses suggested that the left-right coordination swimming pattern is compromised in the absence of Epha4a.

Defects of left-right coordination due to abnormal CPG in the absence of Epha4a

The coordinated left-right locomotion of zebrafish larvae relies on the synchronized contraction of muscle fibers, a process regulated by motor neurons situated on each side of the fish. To explore this intricate mechanism, we utilized a Tg(elavl3:GAL4; UAS: GCaMP) double transgene, allowing the expression of a genetically encoded calcium sensor in all neurons. We observed the rhythmic activation of calcium signaling in motor neurons located within the spinal cord (Fig. 4A, Fig. S6A, Movie S5, Movie S6). In wild-type larvae, the calcium signals exhibited an alternating pattern between the left and right sides of the body, whereas this coordinated pattern was disrupted in both epha4a mutants and morphants (Fig. 4A-B vs A’-B’, Fig. S6A-B vs A’-B’, Movie S7, Movie S8). Additionally, we found that the activation frequency of motor neurons on the left and right sides was comparable in wild-type larvae, but significantly different in the absence of Epha4a (Fig. 4C-D, Fig. S6C-D).

Uncoordinated left-right activation of spinal cord neurons in epha4a mutants.

A, A’: Fluorescent images showing the dorsal-view of 24 hpf Tg(elavl3:GAL4; UAS:GCaMP) double transgenic larvae. The corresponding movies are shown in Movie S5 and S6. B, B’: Line charts showing the quantification of fluorescence changes of the region of interests (ROIs, circled in A, A’) in wild-type larvae and epha4a mutants. C: Scatter plot showing the distribution trend of the ratio of the calcium signal frequency between left and right in wild-type (N=14 larvae, n=42 experiments) and epha4a mutants (N=10 larvae, n=30 experiments). D: Statistical graph of the ratio of the calcium signal frequency between left and right in wild-type larvae and epha4a mutants. Scale bars: 50 μm in panel (A, A’).

One well-established concept regarding left-right coordination involves the presence of central pattern generators (CPGs), which are regulated by interneuron circuitry within the spinal cord. We further examined the axon guidance of interneurons in epha4a mutants. First, we investigated the commissural trajectories of reticulospinal (RS) interneurons. In control larvae, the large Mauthner neurons, along with other RS neurons, were symmetrically positioned on each side of the midline and projected their axons to the contralateral sides (Fig. 5A). These bilaterally projected axons typically crossed at the midline and subsequently synapsed on motoneurons of the opposite sides, contributing to the generation of spinal cord neural circuits (Hale et al., 2016). However, in epha4a mutant larvae, we observed an abnormal pattern in axonal projections. Specifically, the mutant axons failed to traverse the midline and instead extended ipsilaterally (Fig. 5A). In addition, the distance between Mauthner neurons (rhombomere 4) and rhombomere 7 was significantly decreased and rhombomere 5 was scarcely visible in the mutant larvae (Fig. 5A-B). Furthermore, the sites of axon crossing between two Mauthner neurons tended to deviate to one side of the midline in the mutants (Fig. 5C-D’).

Disorganized neural patterning in epha4a mutants.

A: Confocal images showing reticulospinal neuronal axons in 48 hours post-fertilization (hpf) wild-type and epha4a mutant larvae visualized with anti-neurofilament antibody RMO44. Asterisks indicate the cross sites of Mauthner axons. The green arrow indicates the cell body of the r5/MiD2 neuron in a wild-type larva. The red arrow points to the ipsilaterally projected axon of r6/MiD3 in the mutant larva, which is normally projected to the other side in wild-type fish. B: Statistical chart showing the distance between r4 and r7 of 48 hpf wild-type and epha4a mutants. C: Scatter plot showing the distance (d) between the center line and the intersection site of Mauthner axons as indicated in panel C′. D: The ratio of the distance between the intersection site of Mauthner axons and bilateral axon bundles in 48 hpf wild-type (N=11 larvae) and epha4a mutants (N=12 larvae). The ratios were calculated as in panel D′. Scale bars: 20 µm in panel (A).

Cerebrospinal fluid-contacting neurons (CSF-cNs) represent a unique type of interneuron responsible for modulating the V0 and V2a interneurons, which are integral components of locomotor CPGs (Fidelin et al., 2015; Talpalar et al., 2013; Wu et al., 2021). In wild-type larvae, we observed that the ascending axons of these neurons projected either to the right or left side from the midline, as visualized using Tg(pkd2l1:GAL4;UAS:Kaede) double transgenic larvae (Fig. 6A). However, in the absence of Epha4a, the projection of these neurons exhibited notable disorganization, with numerous axons crossing the midline from one side of the trunk. This disorganized pattern was observed in both heterozygotic and homozygotic mutants (Fig. 6A-B). Additionally, we employed an optogenetic approach to activate these CSF-cNs, utilizing the Tg(Gal4s1020t; UAS:ChR2) double transgene. Following optical stimulation, we observed robust tail oscillations as previously described (Wyart et al., 2009). In wild-type larvae, these tail oscillations exhibited a symmetrical left-right beating pattern(Fig. 6C-D, Movie S9). However, a striking disruption of this symmetry was observed in epha4a morphants, as they consistently beat towards one side of the trunk following optical stimulation (a positive angle reflects a rightward turn) (Fig. 6C’-D’ and E, Movie S10). Collectively, these findings demonstrate that the left-right coordination deficiencies observed in epha4a mutants arise from abnormal neural circuit formation, which consequently disrupts the integrity of the CPGs.

Aberrant swimming as a result of abnormal extension of CSF-cNs axons.

A: Fluorescent images showing the distribution of ascending axons of CSF-cNs marked by Tg(pkd2l1:GAL4,UAS:Kaede) in 2 dpf epha4a mutant larvae. Yellow line indicates the midline and the red arrows indicate aberrantly extended axons in epha4a+/- and epha4a-/- larvae. B: Bar graph showing the percentages of abnormal extension of CSF-cNs axons in 2 dpf wild-type, epha4a heterozygote, and epha4a homozygous. The numbers of larvae are indicated in each column. C, C’: Superimposed frames of tail oscillations in 5 dpf control and epha4a morphants. D, D’: A plot of body curvature angles in panel (C) and (C’). The positive angle means turning right. E: Percentages of tendentious swing in control (N=8 larvae, n=24 experiments) and epha4a morphants (N=11 larvae, n=33 experiments). The percentages were calculated by the ratio of tendentious tail oscillation during the first eight swings. Scale bars: 50 µm in panel (A).

Ephrin B3-Epha4 signaling regulates interneuron axon extension

To further explore the role of Epha4a during interneuron axon extension, we examined the expression of epha4a during early zebrafish embryonic development. Whole-mount in situ hybridization results showed that both epha4a and epha4b were abundantly expressed in the zebrafish spinal cord (Fig. S7A). We plotted the expression of these two genes using two published single-cell transcriptome data, which showed that epha4a and epha4b were both expressed in interneurons, suggesting a role for epha4 in interneuron function (Fig. S7B) (Cavone et al., 2021; Scott et al., 2021). Notably, the expression of efnb3b, encoding the ligand of Epha4, was highly enriched in the midline floor plate cells (Fig. S7B).

Ephrins, through interacting with Eph receptors, play a critical role in repulsive axon guidance during neural development (Egea and Klein, 2007; Flanagan and Vanderhaeghen, 1998). We further analyzed axon guidance in efnb3b morphants. Similar to those of epha4a mutants, the efnb3b morphants also displayed axon guidance defects, as well as uncoordinated calcium activation (Fig. S8A-C). In addition, morphants larvae also displayed left-right oscillations defects after optogenetic stimulation (a positive angle reflects a rightward turn) (Fig. S8D-E).

Candidate variants in the EPHA4-related genes

Our zebrafish studies suggested that EPHA4 signaling is crucial for interneuron axon guidance, hence the formation of functional CPGs. Next, we further asked whether mutation of other components of the EPHA4 signaling can result in IS in humans. By searching for rare variants in EPHA4-related genes (Fig. S9) (Szklarczyk et al., 2019), we identified heterozygous de novo start-loss variant (c.1A>G, p.Met1?) in NGEF in a young male (SCO2003P3332) (Fig. 7A-B, Table 1), whose scoliosis was diagnosed in an age range of 15-20, with a main curve Cobb angle of 60°. NGEF encodes the neuronal guanine nucleotide exchange factor Ephexin that differentially affects the activity of GTPases RHOA, RAC1, and CDC42. The activation of Ephexin is triggered by ephrin through EPHA4 (Shamah et al., 2001).

IS patients with potential NGEF variants.

A: Protein structure of NGEF protein with the position of potential variants. B, C: Pedigrees and spinal radiographs of two probands with dominant gene variants. Sanger sequencing results are shown on the right.

Strikingly, we further identified a dominant missense variant (c.857G>A, p.Ala286Val) in NGEF in a quad family with three affected members (Fig. 7C, Table 1). The proband (II:2, TSRHC01) has a 58° main curve Cobb angle. This variant, which is located in the RhoGEF domain of Ephexin, is predicted to be highly deleterious (CADD=29.6).

Altogether, our results suggest that defects of the CPGs owing to abnormal EPHA4 signaling maybe one of the crucial factors responsible for IS.

Discussion

IS is a disease with diverse causes, and the underlying mechanisms can vary even among patients with similar scoliotic phenotypes. Previous genetic studies have highlighted the significance of the extracellular matrix (ECM) in maintaining the balance of axial bone and supporting soft tissues in the spine, thus playing a crucial role in IS development (Haller et al., 2016). For example, the top SNP associated with IS maps to LBX1, an essential molecule for ECM maintenance and bone homeostasis (Takahashi et al., 2011). Additionally, genetic loci in muscle development-related genes have also been associated with the onset of scoliosis, emphasizing the intricate interaction between bones and muscles (Ogura et al., 2015). However, it is worth noting that the variants in these genes explain only a small fraction of the overall heritability of IS. Consequently, it is imperative to establish connections between the extensive genetic findings and biological mechanisms that can elucidate the etiological landscape of IS.

In this study, we mapped 41 significant genome-wide loci to functional genes through positional mapping and functional mapping such as eQTL. Then we determined the enrichment in patient cohorts of rare variants in these genes, which may have greater impacts compared with common SNPs. This approach revealed the convergence of SNPs and rare variants in EPHA4 that are enriched in patients with IS. We also identified additional high-impact variants in NGEF, which is involved in the EPHA4 pathway.

Our studies using zebrafish have revealed that deficiency of Epha4 can lead to the development of scoliosis. Interestingly, we observed that even heterozygotic epha4a zebrafish mutants displayed mild scoliosis (Fig. 2C), which is consistent with the occurrence of this condition in scoliosis patients. Further analysis involving behavior and imaging demonstrated that the absence of Epha4a resulted in defective left-right coordination. This coordination is crucially governed by CPGs, which generate rhythmic patterns of neural activity to coordinate limb movements on both sides of the body (Kiehn, 2006; Marder and Bucher, 2001; Talpalar et al., 2013). Previous studies have reported the involvement of ephrin and its receptors in axon guidance during the maturation of neural circuits, including CPGs (Andersson et al., 2012; Borgius et al., 2014; Iwasato et al., 2007). Ephrin B3, present in radial glial cells at the dorsal midline, serves as a repulsive barrier to axons expressing Epha4 (Imondi et al., 2000). Axon guidance is disrupted when this repulsive pathway is inactivated, leading to midline invasion (Butt et al., 2005; Kullander et al., 2001; Paixão et al., 2013). In Epha4 and Efnb3 knockout mice, a ‘hopping’ gait is observed, which is attributed to the aberrant crossing of motor neurons expressing Epha4 (Kullander et al., 2003). Consistent with these observations, we observed disrupted axon guidance in the interneurons, which are integral components of CPGs, in both epha4a mutants and efnb3b morphants. This finding suggests that CPG malfunction may be a significant contributing factor to the scoliosis phenotype observed in these genetic backgrounds. It is highly likely that the lack of coordinated left-right locomotion generates imbalanced mechanical forces on the spine, gradually leading to spinal curvature during later stages of development (Fig. 8).

The proposed mechanism of IS mediated by EPHA4 dysfunction.

In healthy individuals, EphrinB3-activated EPHA4 phosphorylates CDK5, leading to the phosphorylation of Ephexin, a protein encoded by NGEF. Phosphorylated Ephexin can regulate axon guidance through either activating RHOA or suppressing CDC42 and RAC1 signaling. These processes are critical to maintain the normal function of the CPG, the local neural network that provides coordinated bilateral muscle control. Impairment of the EPHA4 pathway and CPG may cause an imbalance of the motor drive from the spinal cord during development, thus causing the uncoordinated left/right swimming behavior in zebrafish larvae and the asymmetry of the bilateral muscular pull in a young child. Although the appearance is normal in early childhood, the dysfunction produces a scoliotic curve during the growth spurt.

Thus, our data provided a novel biological mechanism of IS, i.e., the impairment of neural patterning and CPG. Previous studies have provided clues on the role of CPGs in IS. Patients with adolescent IS showed asymmetric trunk movement during gait, as characterized by increased relative forward rotation of the right upper body in relation to the pelvis (Kramers-de Quervain et al., 2004; Nishida et al., 2017). An electromyography (EMG) study also showed asymmetric activation of paraspinal muscles between the convex and concave sides at the scoliosis curve apex (Shimode et al., 2003). In a child with a strong family history of IS, asymmetric hyperactivity was observed by EMG months before scoliosis was evident (Valentino et al., 1985). These left-right locomotor coordination abnormalities indicated the maldevelopment of CPGs as potential cause of IS. The CPG asymmetry may induce an imbalance in trunk muscle strength, resulting in asymmetric rib drooping. This leads to an abnormal vertebral rotation and then the onset of IS. Moreover, in a companion study, Wang et al. identified a number of rare variants in SLC6A9, which encodes glycine transporter 1 (GLYT1), in familial and sporadic adolescent IS cases. The slc6a9 mutant zebrafish also exhibited discoordination of spinal neural activities with pronounced lateral spinal curvature, recapitulating the human IS phenotype (Wang et al., 2024). Taken together, we propose that the dysfunction of CPGs would cause an imbalance in the motor drive from the spinal cord and the asymmetric transversospinalis muscle pull, eventually producing a scoliotic curve (Fig. 8).

In summary, our study demonstrates that both common and rare variants within the EPHA4 pathway contribute to the genetic architecture of IS. EPHA4 pathway dysfunction causes axon pathfinding defects, resulting in impaired coordinated left-right locomotion by disrupting neural patterning and the function of CPGs, thereby potentially leading to IS.

Materials and methods

Mapping of candidate genes utilizing previous association studies

Through a systematic literature review, we identified IS-associated SNPs reported in GWASs and meta-analyses of GWAS. The literature search was carried out using MEDLINE (via Pubmed.gov) and Web of Science (via Clarivate Analytics) and was limited to English-language articles published from January 1980 to October 2020. The following keywords were combined to perform the search: ‘idiopathic scoliosis’ AND ‘GWAS’ OR ‘SNP’ OR ‘single nucleotide polymorphism’ OR ‘variant’ (Table S4). The inclusion/exclusion criteria of abstracts are provided in Table S5. After screening the titles and abstracts, we obtained from the full-text articles the rsID, chromosome, and position for SNPs with a threshold for genome-wide significance of P < 5.0 × 10−8. For SNPs identified in multiple studies, we recorded the lowest P-value.

SNPs reported in previous studies were first pruned using Functional Mapping and Annotation (FUMA, v1.4.1, https://fuma.ctglab.nl/) (Watanabe et al., 2017). Significant SNPs were considered independent at r2 < 0.6. All known SNPs (available in the 1000 Genomes reference panel, https://www.internationalgenome.org/) were included for further gene mapping if they were in a linkage disequilibrium (LD) block (r2 ≥ 0.6) with a significant independent SNP. Three SNP-to-gene strategies were used:

For positional mapping of significant independent SNPs, we used annotations obtained from ANNOVAR (http://annovar.openbioinformatics.org). A 10-kb maximum distance was applied for intergenic SNPs.

For the eQTL mapping, we mapped significant independent SNPs and SNPs in an LD block to eQTLs acorss 44 GTEx tissue types (release V8) (Consortium, 2020). SNP-gene pairs with a false discovery rate (FDR) ≤ 0.05 were considered significant.

For chromatin interaction mapping, we overlapped the significant independent SNPs and those in LD blocks with one end of significantly interacting regions across various tissue and cell types. Information on the interacting regions was derived from Hi-C datasets of 21 tissues and cell types provided by GSE87112. The significance of interactions for Hi-C datasets was computed by Fit-Hi-C, with an FDR ≤ 10^-6 considered significant. Genes were mapped if their promoter regions overlapped with another end of the significant interactions. The promoter region was defined as the region from −250 bp to +500 bp relative to the transcription start site.

Cohort description

The Peking Union Medical College Hospital (PUMCH) cohort: The PUMCH cohort comprised 411 unrelated Chinese patients with severe IS (Cobb angle ≥ 40°) who underwent spinal surgery in the PUMCH between October 2017 and March 2022 as part of the Deciphering disorders Involving Scoliosis and COmorbidities (DISCO) study (http://www.discostudy.org/). The clinical diagnosis was confirmed using standing full-spine radiographs, three-dimensional computed tomography, and magnetic resonance imaging. These patients did not show any congenital or neuromuscular defect at the time of recruitment. The control cohort consisted of 3,800 individuals without observable scoliosis and with exome sequencing or genome sequencing performed at PUMCH for clinical or research purposes. Individuals with vertebrae malformation or congenital developmental defects were excluded. A patient with a 2q35-36.2 deletion, including the EPHA4 gene, was also recruited through in-house gene matching of the DISCO study (Li et al., 2015). This patient was subsequently evaluated for scoliosis by examination and radiography.

The East Asian cohort: The East Asian cohort includes totally 6,449 IS patients and 158,068 controls from four independent datasets (Japanese dataset 1 (Kou et al., 2013; Takahashi et al., 2011): 1,261 cases, 15,019 controls; Japanese dataset 2 (Ogura et al., 2015): 878 cases, 21,334 controls; Japanese dataset 3 (Kou et al., 2019): 3,333 cases, 119,630 controls; Hong Kong dataset: 977 cases, 2,085 controls). The inclusion criteria for IS subjects was as same as our previous studies (Fan et al., 2012; Kou et al., 2019).

The Texas Scottish Rite Hospital for Children (TSRHC) cohort: We used a replication cohort of European-ancestry patients with IS from the TSRHC. Cases considered for inclusion in the study met criteria for a positive diagnosis of IS: lateral deviation from the midline greater than 15 degrees as measured by the Cobb angle method from standing spinal radiographs, axial rotation toward the side of the deviation and exclusion of relevant co-existing diagnoses.

Blood sample collection

In the PUMCH cohort, genomic DNA samples were extracted from peripheral blood leukocytes of each subject using a QIAamp DNA Blood Mini Kit (Qiagen, Hilden, Germany), according to the manufacturer’s protocol. Purified DNA was qualified by Nanodrop 2000 (Thermo Fisher Scientific, Waltham, MA, USA) and quantified by Qubit 3.0 using the dsDNA HS Assay Kit (Life Technologies, Carlsbad, CA, USA). DNA samples were stored at 4°C until used.

In the East Asian and TSRHC cohort, genomic DNA was extracted from peripheral blood or saliva using a standard protocol.

DNA sequencing and variant calling

In the PUMCH cohort, whole exome sequencing (WES) or whole genome sequencing (WGS) was performed on peripheral blood DNA from all individuals and available family members (Table S6). A SureSelect Human All Exon V6+UTR r2 core design (91 Mb, Agilent) was used for exon capture. The exomes were then sequenced on an Illumina HiSeq 4000 (Illumina, San Diego, CA, USA) according to the manufacturer’s instructions. For WGS, sequencing libraries were prepared using the KAPA Hyper Prep kit (KAPA Biosystems, Kusatsu, Japan) with an optimized manufacturer’s protocol. We performed multiplex sequencing using an Illumina HiSeq X-Ten sequencer (Illumina, San Diego, CA, USA). Variant calling and annotation were done in-house using the Peking Union Medical College Hospital Pipeline (PUMP) described previously (Chen et al., 2021; Zhao et al., 2021).

In the TSRHC cohort, subjects in 162 adolescent IS families were sequenced as part of the Gabriella Miller Kids First Pediatric Research Consortium (GMKF) at The HudsonAlpha Institute for Biotechnology (Huntsville, AL). In summary, DNA was normalized, sheared and then ligated to Illumina paired-end adaptors. The purified ligated DNA was amplified and exome sequencing was performed on the Illumina HiSeq X platform. The sample’s sequences were aligned into GRCh38 and genotypes were joint-called per each family by the GMKF’s Data Resource Center (DRC) at Children’s Hospital of Philadelphia following GATK best practices as detailed here (Mukhopadhyay et al., 2020). GMKF DRC’s alignment and joint genotyping pipelines are open source and made available to the public via GitHub (https://github.com/kids-first/kf-alignment-workflow and https://github.com/kids-first/kf-jointgenotyping-workflow).

Rare variant association analysis of the candidate genes

To determine the contribution of rare variants in GWAS candidate genes to IS, we analyzed the gene-based mutational burden for 156 candidate genes. Only ultra-rare variants with a gnomAD population-max allele frequency ≤ 0.01% and a cohort allele count ≤ 3 were analyzed. Each variant was assigned a weight, and the mutational burden of a given gene was defined as the maximum weight value among all ultra-rare variants carried by the individual. The SAIGE-GENE+ package was used to determine the weighted mutational burden test for each gene (Zhou et al., 2022).

Weighting criteria for the weighted burden analysis were developed according to the variant types and in silico results. The LoF variants (canonical splicing variants, nonsense variants, or variants that cause frameshift, stop-gain, or start-loss) were calculated together with the protein-altering variants, including missense variants and in-frame indels. Each variant was assigned a weight range from 0-1 and the mutational burden for a given gene was defined as the maximum weight value among all ultra-rare variants carried by the individual. The LoF variants annotated as ‘high confidence’ by the loss-of-function transcript effect estimator (LOFTEE) (Karczewski et al., 2020) were assigned a weight value of 1.0. The LoF variants annotated as ‘low confidence’ or unlabeled by LOFTEE, the non-canonical splicing variants with a SpliceAI score > 0.5, and the missense variants with a rare exome variant ensemble learner (REVEL) (Ioannidis et al., 2016) score of 0.8 were assigned a weight value of 0.8. The in-frame insertions/deletions (indels) with a Combined Annotation Dependent Depletion (CADD) (Kircher et al., 2014) score > 20 and the missense variants with a REVEL score > 0.6 and ≤ 0.8 were assigned a weight value of 0.6. The in-frame indels with a CADD score > 10 and ≤ 20 and the missense variants with a REVEL score > 0.4 and ≤ 0.6 were assigned a weight value of 0.4. The in-frame indels and the missense variants with a REVEL score > 0.2 and ≤ 0.4 were assigned a weight value of 0.2. The remaining missense variants were assigned a weight value of 0.

Sanger sequencing of familial participants

Sanger sequencing of familial participants was performed to determine the origin of variants in EPHA4 and NGEF. All LoF variants and protein-altering variants identified in familial participants were validated. Variant-encoding gene regions were amplified by PCR from genomic DNA obtained from probands, as well as from parents for trios, to determine the origin of the variants. The amplicons were purified using an Axygen AP-GX-50 kit (Corning, NY, USA) and sequenced by Sanger sequencing on an ABI 3730xl instrument (Thermo Fisher Scientific, Waltham, MA, USA).

Minigene assay

The splicing variant (EPHA4: c.1443+1G>C) was characterized by a minigene assay. Genomic DNA from the heterozygous patient was amplified by PCR using a high-fidelity DNA polymerase. Amplicons included exon 6, intron 6, exon 7, intron 7, and exon 8 of the EPHA4 gene. PCR products were cloned into the vector via the restriction sites BamHI and MluI for pCAS2, which is based on the pcDNA3.1 plasmid (Thermo Fisher Scientific, Waltham, MA, USA). Clones with wild-type or mutant genomic inserts were selected and verified by sequencing the cloned DNA fragments. The recombinant plasmids were transfected into 293T cells using LipofectamineTM 3000 reagent (Thermo Fisher Scientific, Waltham, MA, USA). For RT-qPCR, total RNA was isolated from the transfected cells using Trizol reagent (Thermo Fisher Scientific, Waltham, MA, USA and reverse transcription was performed using the GoScript™ Reverse Transcription System (Promega, Madison, MI, USA). PCR amplification was performed using the pCAS2-RT-F and pCAS2-RT-R primers, and the products were sequenced using pCAS2-RT-F.

Primers sequences for the minigene assay were as follows:

pCAS2-RT-F: 5′-CTGACCCTGCTGACCCTCCT-3′

pCAS2-RT-R: 5′-TTGCTGAGAAGGCGTGGTAGAG-3′

Nested PCR

The splicing variant (EPHA4: c.1318+10344A>G) was characterized by nested PCR. RNA was extracted from the whole blood of the patient using the TRIzon Reagent (CWBIO, Hangzhou, China) according to the manufacturer’s guidelines. The cDNA was synthesized using HiScript II 1st Strand cDNA Synthesis Kit with a gDNA wiper (Vazyme, Nanjing, China) according to the manufacturer’s guidelines. Nested PCR was performed as described previously (Yao and Tavis, 2005). Nested PCR reactions in 50 μl, included 2 μl cDNA from the reverse transcription reaction as the template for the first round of PCR or 2 μl of first-round PCR product as the template for the second PCR, 1.5 μl 10 μM sense primer, 1.5 μl 10 μM anti-sense primer, 5 μl nucleotide mix (2 mM each dNTP), 5 μl 10× KOD Buffer, 1 μl 1 unit/μl Kod-Plus-Neo polymerase, 3 μl 25 mM MgSO4 and 31 μl ddH2O. The PCR program was (94°C for 2 min, 98°C for 10 s, 55°C for 30 s, 68°C for 60 s) × 20 cycles and then (94°C for 2 min, 98°C for 10 s, 57°C for 30 s, 68°C for 60 s) × 30 cycles. Primers sequences:

For the first round:

F: 5′-GGCTCCTGTGTCAACAACTC-3′

R: 5′-GTTGGGATCTTCGTACGTAA-3′

For the second round:

F: 5′-AACTGCCTATGCAACGCTGG-3′

R: 5′-AGCTGCAATGAGAATTACC-3′

Western blots

HEK293T cells were transfected with EPHA4-C1-pEGFP plasmid or the corresponding wild type vector to detect the missense variant in EPHA4 (c.2546G>A, p.Cys849Tyr). Cells were cultured in six-well plates and transfected with DNA (2 mg/well) using Lipofectamine 3000 reagent (Thermo Fisher Scientific, Waltham, MA, USA). After 48 hours, cells were harvested, and protein extracts were prepared as described (Ding et al., 2017). The mutant proteins and wild-type proteins were fused to GFP. The expression of the two proteins was compared by western blots with an GFP antibody using the ECL detection system. The GFP antibody was purchased from Cell Signaling Technology (Cell Signaling, Danvers, MA, USA). The CDK5 antibody was purchased from Santa Cruz Biotechnology (Santa Cruz, CA, USA). A phospho-specific antibody against CDK5 phosphorylated at Tyr (GeneTex, Irvine, CA 92606 US) was purchased from GeneTex. Western blot experiments were repeated twice with similar results for the replicates.

Genotyping and imputation of GWAS

In the East Asian Japanese cohort, genotyping was performed by Illumina Human610 Genotyping BeadChip, Illumina HumanOmniExpressExome and HumanOmniExpress as our previous GWASs (Kou et al., 2019; Kou et al., 2013; Ogura et al., 2015; Takahashi et al., 2011). For quality control (QC), subjects with call rate < 0.98, high-degree relatedness with other subjects and outliers of East Asian ethnicity were excluded. For variants QC, the exclusion criteria of variants were as follows: call rate < 0.99, P-value for Hardy-Weinberg equilibrium < 1.0 × 10−6 and minor allele count < 10. The reference panel for imputation, namely JEWEL7K, was composed of 1000 Genomes Project phase 3 (v5) (Auton et al., 2015) and Japanese whole-genome sequence data (Das et al., 2016) with 3256 high-depth subjects (≥30 read counts) and 4216 low-depth subjects (≤ 15 read counts). Using EAGLE 2.4.1 (https://alkesgroup.broadinstitute.org/Eagle/) to determine the haplotypes, pre-phasing was conducted. Genotypes were imputed using Minimac4 (version1.0.0) (Das et al., 2016). After imputation, we excluded variants with minor allele frequency (MAF) < 0.005 and low imputation quality (R2 < 0.3).

In the East Asian Hong Kong cohort, samples were genotyped with Illumina Human Omni ZhongHua-8 Beadchips. Illumina Genome Studio v2.0 was used to convert raw data into PLINK format. The QC steps of samples and variants were described in a previous study (Marees et al., 2018). Genotype phasing and imputing were executed using SHAPEIT v2.r900 (Delaneau et al., 2012) and IMPUTE2 (Marchini et al., 2007). The imputed data was filtered using the following parameters: INFO > 0.6, Certainty > 0.8, and MAF > 0.01. Association analysis was performed using PLINK v1.9 logistic regression with covariates: sex, age and top 20 principal components of the variance-standardized relationship matrix.

East Asian GWAS meta-analysis for IS

For the meta-analysis of the four datasets (three Japanese datasets and one Hong Kong dataset), an inverse-variance-based method was performed by METAL (version2011-03-25) (Willer et al., 2010). SNPs in three or more of the four cohorts were used in subsequent analyses.

Gene-based common variant analysis and eQTL analysis

SNPs in EPHA4 or within 20 kb flanking EPHA4 with a significant association with IS were retrieved. SNPs were matched to potential eQTLs according to the GTEx database (v8, https://gtexportal.org). Gene-based common variant analyses were performed using Multi-marker Analysis of GenoMic Annotation (MAGMA) (de Leeuw et al., 2015) and FUMA (Watanabe et al., 2017) using default settings with LD information from the 1000 Genomes Project East Asian population (1KGP EAS) as a reference. SNPs located within two kb upstream and one kb downstream from EPHA4 were included in the gene-based analysis.

Zebrafish strains, mutants, and morphants

Zebrafish Tuebingen (TU) strains were maintained at 28°C on a 14-hour/10-hour light/dark cycle. Embryos were raised at 28.5°C in E3 medium (5 mM NaCl, 0.17 mM KCl, 0.39 mM CaCl2, 0.67 mM MgSO4) following standard protocols. Zebrafish have two homologs of EPHA4, epha4a and epha4b. The CRISPR/Cas9 system was used to generate zebrafish epha4a and epha4b mutants. To increase efficiency, we injected Cas9 mRNA together with multiple single guide RNAs (sgRNAs) for each gene (Table S7). The sgRNA sequences for epha4a and epha4b are listed in Table S7. Morpholino sequences for epha4a and efnb3b knockdown analysis are also listed in Table S7.

Micro CT imaging

Adult zebrafish epha4a mutants or wild-type siblings were euthanized with tricaine methanesulfonate and fixed in 4% paraformaldehyde. Micro CT images were captured using a PerkinElmer Quantum GX2. Planar images acquired over 360° of rotation were reconstructed using QuantumGX. Three-dimensional renders of the skeleton were made with Analyze 12.0 software (AnalyzeDirect).

Behavioral recordings and analysis

For behavior analysis, individual zebrafish larvae were transferred into 24-well plate with fresh E3 medium at 8 dpf. Then, the plate was placed inside the Daniovision (Noldus) observation chamber for further behavior analysis. Video-tracking of swimming activity and further statistical analysis were performed using the EthoVision™ XT10 software. Behavioral data were shown as total swim distance (mm), average velocity (mm/s), average relative turn angle (degree), average absolute turn angle (degree) and average absolute angular velocity (degree/s) at a duration for 4 min. A positive angle reflects a leftward turn.

For tactile stimulation, zebrafish larvae were placed in a concave slide, and tactile stimulation was performed with glass capillaries to the head or tail sides. Larval startle responses were recorded using a high-speed video camera (Mikrotron, EoSens Mini1) at 1000 fps. Automated analysis of larval movement was performed using the FLOTE software package (Jain et al., 2014).

For optogenetic studies, the Tg(Gal4s1020t; UAS:ChR2) transgenic embryos were injected with control, epha4a or efnb3b morpholinos at one cell stage. At 5 dpf, the head of injected larva was mounted in 1% low melting-point agarose (Sigma) and the body was exposed in glass-bottom dishes (WPI). Leica M165FC fluorescence microscope was used to irradiate with a blue laser at 488nm wavelength. Larval responses were recorded using a high-speed video camera (Mikrotron, EoSens Mini1) at 500 fps. Body curvature analysis of larval movement was performed using the FLOTE software package.

Whole-mount in situ hybridization and immunofluorescence

The primer sequences used to amplify epha4a, epha4b and rfng genes were listed in Table S7. Probe synthesis and whole-mount in situ hybridization were performed according to standard protocols. For reticulospinal neurons immunostaining, embryos were fixed in 2% trichloroacetic acid at 48 hpf for 3-4 h, washed twice in 0.5% Triton X-100 in PBS and blocked in 0.5% Triton X-100, 10% normal goat serum, 0.1% bovine serum albumin (Solarbio) in PBS for 1 h. The embryos were stained by monoclonal anti-neurofilament 160 antibody (Sigma-Aldrich) overnight at 4°C, then stained by goat anti-mouse Alexa Fluor 488 (Invitrogen) after washing.

Analysis of neuronal calcium signals activity

Control, epha4a or efnb3b morpholinos were injected into Tg(elavl3:GAL4; UAS:GCaMP) transgenic embryos at one cell stage. At 24 hpf, embryos were paralyzed with 0.5 mg/ml α-Bungarotoxin (AlomoneLabs), then mounted in 1% low melting-point agarose (Sigma) in glass-bottom dishes (WPI). Neuronal calcium signal images were collected with IXON-L-888 EMCCD camera equipped on Dragonfly 200 Spinning Disk Confocal Microscope using a 20×/0.55 objective within 1 min at a frame rate of 10 fps.

Images were analyzed with imageJ software. To quantify the change in fluorescence intensity, a region of interest (ROI) was defined, and the fluorescence intensity Ft of different frames was normalized to Ft/Fmin based on the minimum fluorescence intensity Fmin in all frames. To compare the left-right alternation pattern of neuronal calcium signals, the ratio of the left-side calcium signals frequency (fleft) to the right-side calcium signals frequency (fright) was logarithmically transformed (lg-transformed).

Quantification and statistical analysis

Statistical analyses were performed in SPSS (version 15.0). Unpaired Student’s t-test, Welch one-way ANOVA or two-way ANOVA followed by Turkey’s multiple comparison test were applied when appropriate. All experiments were replicated at least three times independently. P<0.05 was considered statistically significant.

Study approval

Approval for the study was obtained from the ethics committee at the Peking Union Medical College Hospital (JS-098, JS-2364), the medical ethics committee of the Keio University Hospital (No. 20080129), the ethical committee of RIKEN Yokohama Institute (No. H20-17(8)), and the Institutional Review Board of the University Texas Southwestern Medical Center (protocol STU 112010L150). Written informed consent was obtained from each participating individuals and families in the three cohorts. For the control group, the protocols were approved by the ethics committee at Peking Union Medical College Hospital. All zebrafish studies were approved by the Animal Care Committee of the Ocean University of China (Animal protocol number: OUC2012316).